### 2019-01-12 notes-duality/bo-notes.tex
bo-notes.tex

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amsfonts,amssymb}

\newtheorem{theorem}{Theorem}
\newtheorem{claim}{Claim}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\DeclareMathOperator*{\argmin}{\textrm{argmin}}
\DeclareMathOperator*{\argmax}{\textrm{argmax}}
\DeclareMathOperator*{\E}{\mathbb{E}}

\newcommand{\reals}{\mathbb{R}}

\begin{document}

\begin{definition}
  Given a polytope $\{p \in \Delta_n : B p \geq \vec{0}\}$ where $B \in \reals^{k \times n}$, a \emph{$d$-representation} is a collection of $k$ unit vectors in $\reals^d$ satisfying the following.
  Forming the matrix $V \in \reals^{k \times d}$ whose rows are the unit vectors, we must have for each $y$ that the polytope $\{ w : Vw \leq B_y \}$ is nonempty for each $y$ (where $B_y$ is the $y$th column).
\end{definition}

\begin{theorem}
  If an elicitable finite property is $d$-embeddable, then each of its level sets has a $d$-representation.
\end{theorem}
\begin{proof}
  Let $\ell_y: \reals^d \to \reals$ be the $y$th loss function.
  Write $\ell(r;p) = \sum_y p_y \ell_y(r)$ for the expected loss under report $r$.
  In particular, take $r$ to be the embedding point for a given level set.
  In this case, $r$ is optimal for belief $p$ if and only if $0 \in \partial \ell(r;p)$ where the subgradients are taken with respect to $r$ only.
  Therefore, the level set is exactly those $p$ such that $0 \in \partial \ell(r;p)$.
  We now construct these constraints from the loss functions.

  Because $\ell_y$ is a polyhedral convex function, its subgradient set $\partial \ell_y(r)$ is a polytope and can be described as follows.
  For any vector $v \in \reals^d$, the \emph{face exposed by $v$} is the set $F_{vy} = \argmax_{w \in \partial \ell_y(r)} \langle v, w \rangle$ with corresponding maximum value $b_{vy} = \langle v, w \rangle$ where $w \in F_{vy}$.
  Observe that for the polytope $\alpha \partial \ell_y(r)$ for positive $\alpha$, the vector $v$ exposes the face $\alpha F_{vy}$ with maximum value $\alpha b_{vy}$.

  Now, the subgradient set of a weighted sum is the weighted Minkowski sum of the subgradient sets:
    \[ \partial \ell(r;p) = \sum_y p_y \partial \ell_y(r) \]
  It follows that $\partial \ell(r;p)$ is the set of points $w$ satisfying that for all $v$, $\langle v,w \rangle \leq \sum_y p_y b_{vy}$.
  In particular therefore, $0 \in \partial \ell(r;p)$ if and only if $p$ is in the level set if and only if $0 \leq \sum_y p_y b_{vy}$ for all $v$.
  It is a polytope, so we can collect a finite set of $k$ vectors $v_1,\dots,v_k$ and form a matrix $B$ where $B_{iy} = b_{v_iy}$, and the level set is $\{p : B p \geq 0 \}$.
  Collecting $v_1,\dots,v_k$ as the rows of a matrix $V$, it is a $d$-representation of the level set, since the polytopes are nonempty, in particular:
    \[ \partial \ell_y(r) \subseteq \{ w : \langle v_i, w \rangle \leq f_{vy} ~ (\forall i) \}. \]
\end{proof}

We use this to prove lower bounds as follows:
\begin{enumerate}
  \item A level set is given to us as $\{ p : Bp \geq \vec{0}\}$ for a matrix $B \in \reals^{k \times n}$.
  \item We show there is no set of $k$ vectors $v_1,\dots,v_k$ that can form a $d$-representation.
  \item In particular, we show that however you pick $v_1,\dots,v_k$, for one of the outcomes $y$, the polytope $\{w : Vw \leq (B_{1y},\dots,B_{ky}) \}$ is empty (infeasible).
\end{enumerate}

\end{document}