### 2020-01-07 notes/polytope-notes.tex
Pass over by Jessie, making notation consistent.

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{hyperref}       % hyperlinks  %[implicit=false, bookmarks=false]
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage[margin=1.0in]{geometry}

\usepackage{mathtools, amsmath, amssymb, amsthm, graphicx, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}

\usetikzlibrary{calc}
\newcommand{\Comments}{1}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
  \todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{\mytodo{orange!20!white}{BTW: #1}}
\ifnum\Comments=1               % fix margins for todonotes
  \setlength{\marginparwidth}{1in}
\fi

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\dom}{\mathrm{dom}}
\newcommand{\epi}{\text{epi}}
\newcommand{\relint}{\mathrm{relint}}
\newcommand{\prop}[1]{\Gamma[#1]}
\newcommand{\eliccts}{\mathrm{elic}_\mathrm{cts}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}

\newcommand{\cell}{\mathrm{cell}}

\newcommand{\abstain}[1]{\mathrm{abstain}_{#1}}
\newcommand{\mode}{\mathrm{mode}}

\newcommand{\simplex}{\Delta_\Y}

% alphabetical order, by convention
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\F}{\mathcal{F}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\renewcommand{\P}{\mathcal{P}}

\newcommand{\risk}[1]{\underline{#1}}
\newcommand{\inprod}[2]{\langle #1, #2 \rangle}%\mathrm{int}(#1)}
\newcommand{\inter}[1]{\mathrm{int}(#1)}%\mathrm{int}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\indopp}{\bar{\mathbbm{1}}}
\newcommand{\ones}{\mathbbm{1}}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

\newcommand{\Ind}[1]{\mathbf{1}\{#1\}}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\arginf}{arg\,inf}
\DeclareMathOperator*{\sgn}{sgn}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{claim}{Claim}

\title{Polytope notes}
\date{}

\begin{document}
\maketitle

\section{Preliminaries}
\begin{definition}
	A \emph{polyhedra} in $\reals^d$ is defined by the intersection of a finite number of half-spaces.
	A \emph{polytope} is a bounded polyhedra.
\end{definition}

\begin{definition}[Valid inequality]
	Let $S$ be a set in $\reals^d$.
	A \emph{valid inequality} for $S$ is an inequality that holds for all vectors in $S$.
	That is, the pair $(a,\beta)$ is a valid inequality for $S$ if and only if 
	\begin{align*}
	\inprod{a}{x} &\leq \beta \; \; \forall x \in S~.~
	\end{align*}
\end{definition}

\begin{definition}[Face]\label{def:face}
	For any valid inequality of a polytope, the subset of the polytope of vectors which are tight for the inequality is called a \emph{face} of the polytope.
	That is, the set $F$ is a face of the polytope $T$ if and only if 
	\begin{align*}
	F &= \{x \in T : \inprod{a}{x} = \beta \}
	\end{align*}
	for some valid inequality $(a, \beta)$ of $T$.
\end{definition}

\begin{definition}[Supporting function]
	Let $S$ be a nonempty bounded set in $\reals^d$.
	We call the \emph{supporting function} of $S$ the function $H_S:\reals^d \to \reals$ by
	\begin{align*}
	H_S(a) := \sup_{x \in S}\inprod{a}{x}~.~
	\end{align*} 
\end{definition}

\begin{definition}[Maximizers]
  Let $S \subseteq \reals^d$, and $a \in \reals^d$.
  The \emph{set of maximizers} of $a$ over $S$ is defined as
  \begin{align*}
    \mathcal{S}(S;a) &= \{x \in S : \inprod a x = H_S(a)\}
  \end{align*}
\end{definition}

\begin{definition}[Normal cones]
  Let $T$ be a polytope in $\reals^d$.
  For any face $F$ of $T$, we define its \emph{normal cone} $\N(F;T)$ as the set of vectors for which $F$ is the maximizer set over $T$.
  That is,
  \begin{align*}
    \N(F;T) = \left\{a : F = \mathcal{S}(T; a) \right\}~.~
  \end{align*}
\end{definition}

  It is worth noting that normal cones are generally not closed by this definition, but sometimes we may want to think about the closure of the normal cone.

\begin{definition}[Minkowski sum]
	Let $S_1, S_2, \ldots, S_n$ be sets of vectors.
	We can define their \emph{Minkowski sum} as the set of vectors which can be written as the sum of a vector in each set.
	Namely,
	\begin{align*}
	S_1 \oplus \ldots \oplus S_n &= \{x_1 + \ldots + x_n : x_i \in S_i \; \forall i \}
	\end{align*}
\end{definition}

\begin{theorem}[EPFL Thesis Theorem 3.1.2]\label{thm:unique-face-decomp}
	Let $T_1, \ldots, T_n$ be polytopes in $\reals^d$ and let $F$ be a face of the Minkowski sum $T := T_1 \oplus \ldots \oplus T_n$.
	Then there are faces $F_1, \ldots, F_n$ of $T_1, \ldots, T_n$ respectively such that $F = F_1 \oplus \ldots \oplus F_n$.
	Moreover, this decomposition is unique.
\end{theorem}

\begin{corollary}[EPFL Cor 3.1.3]\label{cor:face-decomp-normal-cones}
  Let $T = T_1 \oplus \ldots \oplus T_n$ be a Minkowski sum of polytopes in $\reals^d$, let $F$ be a nonempty face of $T$ , and let $F_1, \ldots, F_n$ be its decomposition.
  Then $\N(F;T) = \N(F_1;T_1) \cap \ldots \cap \N(F_n; T_n)$.
\end{corollary}

\begin{corollary}[EPFL Cor 3.1.4]
  Let $F_1, \ldots, F_n$ be nonempty faces of the polytopes $T_1, \ldots, T_n$ respectively, then $F_1 \oplus \ldots \oplus F_n$ is a face of $T_1 \oplus \ldots \oplus T_n$ if and only if the intersection of their normal cones is nonempty.
\end{corollary}

\begin{theorem}[EPFL Th 3.1.6]\label{thm:support-minksum}
  The supporting function of a Minkowski sum is the sum of the supporting functions of its summands.
\end{theorem}

\begin{definition}[Weighted Minkowski sum]
  If $T_1, \ldots, T_n$ are polytopes in $\reals^d$, we can call $T(\vec p)$ their \emph{weighted} Minkowski sum for $\vec p \in \reals^n_+$
  \begin{align*}
    T(\vec p) &:= \oplus_y p_y T_y = p_1 T_1 \oplus \ldots \oplus p_n T_n
  \end{align*}
\end{definition}

  From the thesis:  \emph{``It is easy to see that the normal fan (undefined here, but consequently normal cones) of $p_i T_i$ does not change as long as $p_i$ is positive.  Since the normal fan of a Minkowski sum can be deduced from that of its summands, we can deduce from this that the conbimatorial properties of $\oplus_y p_y T_y$ stay the same as long as all $p_i$ are positive.''}

  \section{Starting with polytopes}\label{sec:start-polytope}
  In this section, we will suppose we start with a finite set of $n$ polytopes $\T := \{T_1, \ldots, T_n\}$, and we will call $T := T_1 \oplus \ldots \oplus T_n \in \reals^d$ their Minkowski sum.
  We know that every polytope has both a halfspace and vertex representation ($\H$-representation and $\V$-representation, respectively.)
  In order to construct the $\H$-representation, we know there must be a matrix $V \in \reals^{k \times d}$ and vector $e \in \reals^k$ such that $T = \{x \in \reals^d : Vx \leq e\}$, but it is unclear how to find such a set, which we discuss in Section~\ref{sec:find-V}.

  \subsection{Notation and new definitions}
  We will use the following notation throughout the rest of this section.
  Suppose we are given a polytope $T_y \in \reals^d$ and set of vectors $V \in \reals^{k \times d}$.
  Call $e^y \in \reals^k$ the vector such that $e^y_i = \max_{x \in T_y}\inprod{v_i}{x}$.  
  For a finite set $\T = \{T_1, , \ldots, T_n\}$, let us denote the matrix $E = (e^y)_{y=1}^n$.
  \begin{definition}
    We say a set of normals $V$ is \emph{complete} with respect to a polytope $T_y$ if $T_y = \{x \in \reals^d: Vx \leq e^y\}$.
  \end{definition}
  Moreover, we say $V$ is complete with respect to the set of polytopes $\T$ if and only if $V$ is complete with respect to each $T_y \in \T$.

  \subsection{Finding $V$}\label{sec:find-V}
  \btw{Not sure if we want to go this far, or just argue that a complete set of normals exists.  This would be where Corollaries 1 and 2 come in though.}
  In order to find a matrix of normals $V$ that is complete with respect to $T := \oplus_y T_y$, we can look at the normal cones for each $T_y$ and find a vector in the normal cone $\N(F_y, T_y)$ in the decomposition of $F$ in $T$ for each face $F$.
  Since we know $T_1, \ldots, T_n$ are polytopes and therefore have $\H$-representations, one of the first natural steps is to find a set of valid inequalities describing each $T_y$.
  For each face $F_j$ of the Minkowski sum polytope 
  $T$, we can choose a normal $v_j \in \N(F_j; T)$ and describe the face $F$ by $\{x \in T:  \inprod{v_j}{x} = \sup_{x \in T} \inprod{v_j}{x} =: e_j\}$ by Theorem~\ref{thm:unique-face-decomp} and Definition~\ref{def:face}.
  For each face $F_{j}^{y}$ that is in the decomposition of the face $F_j$, we have $\inprod {v_j}{x} \leq e^y_j$ forming a tight bound on $F_j$ by Theorem~\ref{thm:support-minksum}.

  For ease of notation, we concatenate the $e^y$ vectors corresponding to the support of each normal defining the polytope $T_y$ to form the matrix $E$ according to a consistent ordering on $\Y$.
  Note that as we assume $V$ is complete with respect to $\T$, we have $e^y$ having the same dimensions for all $y$, and thus $E$ is a matrix.

%  Therefore, if we take a matrix $V$ that is a finite set of the elements of the normal cones for each face in $T$, such a matrix can also be used to define the $\H$-representation for not only $T$, but each $T_y$.
%  For each face $F_i$ of $T$ with $\N(F_i, T) \neq \emptyset$, we can choose one normal $v_i$ so that the face $F_i = \{x : \inprod{v_i}{x} = \sup_{x\in T}\inprod{v_i}{x} =: e_i\}$.
%  As this happens for each face with a nonempty normal cone, we can repeat this for each $T_y \in \T$.
%  This yields the matrix $E := (e_{yj})$ where $e_{yj} = \sup_{x \in T_y}\inprod{v_j}{x}$.
%

  \begin{proposition}\label{prop:complete-V}
  	The normals $V$ constructed by taking an element of the normal cone $\N(F, T)$ for each face $F$ in $T$ is complete for both $T$ and each $T_y \in \T$.
  \end{proposition}
  \begin{proof}
  	First, to see this constructed $V$ is complete for $T$, consider that the normals to the facets (in the affine span of $T$) of $T$ is complete for $T$.
  	Moreover, each additional normal is a redundant constraint, so
%   since the inner product of elements of the normal cone is less than or equal to the support for the polytope for every element of $T$.
%  	Thus, 
  	$V$ is complete for $T$.

  	To see why $V$ is complete with respect to each $T_y$, consider each case of subset inclusion.
  	We will let $\bar T_y := \{ x : Vx \leq e^y\}$ be the polytope described by $V$ and $E$, and $T_y$ be the given polytope.

  	First,  to see $T_y \subseteq \bar T_y$, take $x \in T_y$.
  	For all $i \in [k]$, we then have $\inprod{v_i}{x} \leq H_{T_y}(v_i) = e^y_i$.
  	By definition, this holds for all $v_i$ such that $v_i \in \N(F, T_y)$ for some $F$ a face of $T_y$.
  	We then want to claim that, for all $v_i \in V$, there is some face $F_y$ of $T_y$ such that $v_i \in \N(F_y, T_y)$, and thus the previous statement holds.
  	This follows from Corollary~\ref{cor:face-decomp-normal-cones}, since we know that we must have $v_i \in \N(F, T)$ for some face $F$ of $T$ by construction.
  	We necessarily have $v_i \in \N(F_y, T_y)$, where $F_y$ is the face of $T_y$ in the decomposition of $F$.

  	Now to see $\bar T_y \subseteq T_y$, we use a contradiction.
  	Suppose there was some $x \in \bar T_y$ such that $x \not \in T_y$.
  	Then there must be some $v_i \in V$ so that $\inprod {v_i}{x} > \sup_{y \in T_y} \inprod{v_i}{y}$.
  	This implies that for all faces $F_{j}^y$ of $T_y$, we have $v_i \not \in \N(F_{j}^y, T_y)$.
  	However, by construction of $V$, we can consider the face $F$ of $T$ such that $v_i \in \N(F,T) = \cap_y \N(F_y, T_y)$ by Corollary~\ref{cor:face-decomp-normal-cones}.
  	We necessarily have $v_i \in \N(F_y, T_y)$, yielding a contradiction.
  	Thus, we have $v_k \in \N(F_y, T_y)$, so we conclude $\bar T_y \subseteq T_y$, yielding the equality of each $\bar T_y = T_y$, which is the statement.
  \end{proof}

  The construction of $V$ in Proposition~\ref{prop:complete-V} is typically not the minimal $V$ that is complete for $\T$.
  By minimal, we mean that the number of normals $k$ is minimized while still being complete.
  It is unclear at the current moment how to find such a minimal set of normals, so we proceed to understanding what we can say once we have $V$.

  \subsection{Once we have $V$}
  Now, for a given polytope $T(p)$, we will later want to ask when a given $z \in \reals^d$ is in the polytope $T(p)$.
  We will then use this to understand the set of $p \in \simplex$ for which $z \in T(p)$, later substituting $z=0$.
  Throughout, assume we have $V$ which is complete for $\T$ and $E$ defined by the support of each normal in $V$ for all $T_y \in \T$.
  As a reminder, we denote $e^y = E_{;y}$ as the $y^th$ column of $E$, or equivalently, the support vector for $T_y$ given $V$.

  Since we define $T_y = \{x : Vx \leq e^y\}$, we can multiply the right side of the inequality by the constant $p_y \geq  0$ to yield $p_y T_y = \{x : Vx \leq p_y e^y\}$.
  Now in taking the Minkowski sum of polytopes described by the same set of normals, we can take 
  \begin{align*}
  \oplus_y p_y T_y &= \{x : Vx \leq p_1 E_{;1}\} \oplus \ldots \oplus \{x : Vx \leq p_n E_{;n}\} \\
  &= \{x : Vx \leq p_1 E_{;1} + \ldots + p_n E_{;n}\}\\
  &= \{x : Vx \leq E p\}~.~
  \end{align*}
  The first to second line follows from Theorem~\ref{thm:support-minksum} and preservation of inequalities under addition.
  Now, we have $z \in T(p) \iff \inprod{v_i}{z} \leq (Ep)_i$ for all $v_i \in V$.

  We are particularly interested in when $\vec 0 \in T(p)$, we have $\vec 0 \in T(p)$ if and only if $E p \geq 0$ by substitution.  \btw{For later, common result to say $T(p) = \partial \inprod p {L(\cdot)}$, so if $0 \in T(p)$, we have a minimizer of the loss.}

  We assume $p \in \simplex$, so we now describe the cell $D^\T := \{p \in \simplex : Ep \geq \vec 0\}$ as the set of distributions such that $\vec 0 \in T(p)$.
  In cases where $\T$ is obvious from context, we omit the subscript and just write $D$.

  Given the complete set of normals $V$ and constructing the support matrix for $V$ and $\T$, $E$, we observe that $E$ is unique up to rescaling.
  However, as discussed earlier, there are always multiple complete sets of normals for $\T$, and so in that sense, $E$ is not unique.

  \section{Starting at the cell $C$}\label{sec:start-cell}

  Now let us consider some cell $C \subseteq \simplex$.
  Let $A$ be the affine span of the simplex, $A := \{p \in \reals^n : \sum_y p_y = 1\}$.
  We can represent the cell $C := \{p \in A : Bp \geq c\}$.
  In fact, we can set $c := \vec 0$ in order to enforce a unique representation of $B$, up to row reordering and scaling.
  This also allows us to use the same matrix for the $\H$-representation even as we scale the simplex by a positive constant.

  Here, unlike the matrix $E$ when given $V$, the matrix $B$ is not unique.
  In a similar manner, we can simply add redundant constraints in $A$ and yield the same cell $C$.

  \subsection{Notes post NeurIPS} 
  Still not sure if we can construct $B$ with the assumption that $p \in \simplex$ rather than $p \in A$, but gathering from our conversations at NeurIPS, that question seems worth investigating, just maybe after COLT.
  I mean that in the sense that this $B$ matrix captures our necessary condition.
  The vertex representation seems to capture our sufficient condition, possibly rendering the simplex halfspace constraints unnecessary when we think of them as necessary conditions.
  With the results below (that nonsimplex constraints must appear in both $E$ and $B$), being able to assume the simplex for $B$ would simplify a bit... but doesn't give us a polytope construction by any means.

  \section{Equivalences; moving from polytopes to cells and back}

  In the big picture, we want to know two things.
  First, if we are given a cell $C$, how we can come up with a finite set of polytopes $\T = \{T_1, \ldots, T_n\}$ such that $\vec 0 \in T(p) \iff p \in C$.
  Second, we want to know the opposite case: starting from $\T$, can we derive the cell $C \subseteq \simplex$ where $\vec 0 \in T(p)$ for all $p \in C$?

  We know that if we are given $\T$ and a complete set of normals $V$, we can describe $D = \{p \in \simplex : Ep \geq \vec 0\}$ as in Section~\ref{sec:start-polytope}.
  However, we do not necessarily have $D' := \{p \in A : Ep \geq \vec 0\} \subseteq \simplex$, and this poses some issues given the construction of the cell $C$ only assuming that we are in the affine span of the simplex, but not necessarily in the simplex.

  \begin{lemma}\label{lem:describe-D}
    Suppose we have polytopes $\T = \{T_1, \ldots, T_n\}$ and a set of normals $V$ that is complete for $\T$.
 	Take $E = (e_{i}^y)$ where $e_{i}^y = \max_{x \in T_y} \inprod{v_i}{x}$, and $D = \{p \in \simplex : Ep \geq \vec 0\}$.

 	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = \{p \in \simplex: Ep \geq \vec 0\}$.\btw{on the whiteboard: ``stuff about $E$ and $D$.''}
  \end{lemma}
  \begin{proof}
    First, let us fix a distribution $p \in \simplex$.
    %We define $T(p) := \oplus_y p_y T_y$ to be the $p$-weighted Minkowski sum over the polytopes in $\T$.
    By Theorem~\ref{thm:support-minksum}, we have the support of the (weighted) Minkowski sum is the (weighted) sum of the support of each polytope, which we can re-write the weighted support as the product $Ep$.

    Each halfspace is bounded by the support function of the weighted polytope by construction of $E$, so the support of the weighted polytope defined by an inequality on $v_i$ can be described as $\inprod{v_i}{z} \leq \inprod{E_i}{p}$.
    Taking this for all $v_i$, we then have $T(p) = \oplus_y p_y T_y = \{x \in \reals^d : Vx \leq Ep\}$.

    Therefore, for fixed $p$, we have $\vec 0 \in \oplus_y p_y T_y \iff Ep \geq \vec 0$.
    Since this is true for all $p \in \simplex$, we observe the stated set equality.
  \end{proof}

  \begin{proposition}\label{prop:relate-E-B}
    Suppose we have polytopes $\T = \{T_1, \ldots, T_n\}$ and a set of normals $V$ that is complete for $\T$.
    Take $E = (e_{iy})$ where $e_{iy} = \max_{x \in T_y} \inprod{v_i}{x}$, and take $D = \{p \in \simplex : Ep \geq \vec 0\}$ and $C = \{p \in A : Bp \geq \vec 0\}$.

    Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = C$	if and only if $C = D$.
    \btw{whiteboard: ``stuff about $E$ and $B$.''}
  \end{proposition}
  \begin{proof}
  	By Lemma~\ref{lem:describe-D}, we have $D = \{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\}$ by construction, so the result follows from substitution.
  \end{proof}

  \begin{definition}
  	We say a vector $v$ is \emph{redundant} with respect to matrix $Y$ if we have $\{z \in A : Yz \geq \vec 0\} = \{z \in A : [Y;v]z \geq \vec 0\}$.
  \end{definition}

  \begin{proposition}\label{prop:relate-rows}
  	Suppose we have polytopes $\T = \{T_1, \ldots, T_n\}$ and a set of normals $V$ that is complete for $\T$.
  	Take $E = (e_{i}^y)$ where $e_{i}^y = \max_{x \in T_y} \inprod{v_i}{x}$, and take $D = \{p \in \simplex : Ep \geq \vec 0\}$ and $C = \{p \in A : Bp \geq \vec 0\}$.

  	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = C$	if and only if non-simplex rows of $B$ appear in $E$ (possibly scaled) and every other row of $E$ is redundant with respect to $B$.
  	\btw{``replace prop 1 statement with `non-simplex rows of $B$ appear in $E$ (possible scaled) and every other row of $E$ is redundant with respect to $B$.'''}
  \end{proposition}
  \begin{proof}
	\begin{itemize}
		\item [$\implies$] First, assume $C = \{p \in \simplex: \vec 0 \in \oplus_y p_y T_y\}$.
		By Proposition~\ref{prop:relate-E-B}, we know that $C = D := \{p \in \simplex : Ep \geq \vec 0\}$.
		We can re-write $D$ as $\{p \in A : [E;I_n] p \geq \vec 0\}$, building simplex constraints into the inequality rather than the domain assumption.
		By the assumed equality, every row of $B$ must then appear in $[E;I_n]$, where $I_n$ is the $n$-element identity matrix.
		As $\{p \in A : I_n p \geq \vec 0\}$ strictly defines the simplex, any non-simplex row of $B$ must then be described by a row of $E$, up to rescaling.

		What's more is that since the polytopes are the same, they have a unique minimal $\H$-representation (up to rescaling) since they are defined by an inequality on $\vec 0$.
		\btw{Elaborated this justification in the comment below ``As $D$ is defined... yielding a contradiction.''}
%		As $D$ is defined by distributions on $\simplex$ and $C$ on $A$, and $\simplex \subset A$, we have the rest of the rows in $E$ necessarily being redundant with respect to $B$.
%		To see this, suppose there was a row of $E$ that was not redundant with respect to $B$.
%		If such a row did not partition the simplex, it would necessarily be redundant by the restriction that $C \subseteq \simplex$.
%		Thus, such a row much partition the simplex, but not appear in $B$ up to rescaling.
%		Again, if the constraint does not partition the cell, the constraint is redundant, and if it does, then $C \neq D$, yielding a contradiction.

		\item [$\impliedby$] Suppose that any non-simplex rows of $B$ appear in $E$, and every other row of $E$ is redundant with respect to $B$.
		Then we have $D = \{p\in \simplex : Ep \geq \vec 0\} = \{p \in \simplex : Bp \geq \vec 0\}$.
		This polytope is equal to $\{p \in A : Bp \geq \vec 0\} = C$ since we require $C\subseteq \simplex$, so $B$ encapsulates any non-redundant simplex constraints.

		Then $D = C$, and therefore, by Proposition~\ref{prop:relate-E-B}, we have $C = \{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\}$.
	\end{itemize}
  \end{proof}

%  \bigskip
%  \hrule
%  \bigskip
%
%  \section{Necessary and sufficient conditions}
%  \btw{This probably shouldn't be in this document... let's move it later.}
%  \subsection{Necessary conditions}
%    Halfspace representation
%  \subsection{Sufficient conditions}
%    Vertex representation

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: