### 2020-06-02 neurips-20/tikz/t-example.tex
Fixed example in appendix

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usetikzlibrary{calc}
\newcommand{\R}{\mathcal{R}}

\begin{document}

%t example
\begin{tikzpicture}
\draw[opacity = 0.2] (2,0) -- (-2,0) -- (0,3.46) -- cycle;
%label outcomes
\node at (-2, -0.25) {rainy};
\node at (1/2, 3.5) {sunny};
\node at (2, -0.25) {snowy};
%level sets

\draw[fill = blue, fill opacity = 0.1, opacity = 0.2] (0, 3.46) -- (-0.5, 2.595) -- (0.5, 2.595) -- cycle; 
\node at (0, 2.75) {\footnotesize sunny};

\draw[fill = yellow, fill opacity = 0.1, opacity = 0.2] (-2,0) -- (-0.5, 2.595) -- (0, 2.595) -- (0,0) -- cycle; \node at (-0.7, 1) {\footnotesize rainy};
\draw[fill = green, fill opacity = 0.1, opacity = 0.2] (2,0) -- (0.5, 2.595) -- (0, 2.595) -- (0,0) -- cycle; \node at (0.7, 1.) {\footnotesize snowy};

\node at (0,2.595) {\textbullet};

\end{tikzpicture}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: