### 2021-05-11 rates-neurips-21/appendix.tex
raf and bo rates paper

\appendix

\section{Omitted Jokes}

Q: why is a surrogate loss function like a popup ad?

A: minimizing either is inevitably futile.

\vskip1em
Q: Why is walking into a closed door more painful than one that is slightly ajar?

A: hinge loss.

\vskip1em
Q: what do you call a loss function whose penalties are drawn uniformly at random?

A: peer review.