### 2021-05-28 rates-neurips-21/rates.tex
done with intro, very sketchy start to related work

\documentclass{article}

%\usepackage[preprint]{neurips_2021}  % for arxiv
%\usepackage[final]{neurips_2021}     % camera ready
\usepackage[nonatbib]{neurips_2021}             % USE THIS for paper submission

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\usepackage[numbers, sort, compress]{natbib}

\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}

\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
\usepackage{bbm}
\usepackage{amsmath,amsthm,amsfonts,amssymb}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{fact}{Fact}
\newtheorem{observation}{Observation}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

\newtheorem*{theorem*}{Theorem}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{proposition*}{Proposition}

\theoremstyle{definition}\newtheorem{definition}{Definition}
\theoremstyle{definition}\newtheorem{assumption}{Assumption}

\newcommand{\Comments}{1}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
  \todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\proposedadd}[1]{\mynote{orange}{#1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{\mytodo{gray!20!white}{BTW: #1}}%TURN OFF FOR NOW \mytodo{gray}{#1}}
%\ifnum\Comments=1               % fix margins for todonotes
%  \setlength{\marginparwidth}{1in}
%\fi

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\arginf}{arg\,inf}

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\dom}{\mathrm{dom}}
\newcommand{\defeq}{\doteq}%\vcentcolon=} % define equals

\newcommand{\prop}[1]{\mathrm{prop}[#1]}

\newcommand{\simplex}{\Delta_\Y}

% alphabetical order, by convention
\newcommand{\A}{\mathcal{A}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\DeclareMathOperator{\E}{\mathbb{E}}  % expectation
%\newcommand{\E}{\mathbb{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Sc}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}

\newcommand{\risk}[1]{\underline{#1}}
\newcommand{\inprod}[2]{\langle #1, #2 \rangle}%\mathrm{int}(#1)}
\newcommand{\inter}[1]{\mathring{#1}}%\mathrm{int}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\indopp}{\bar{\mathbbm{1}}}
\newcommand{\ones}{\mathbbm{1}}
% \DeclarePairedDelimiter\ceil{\lceil}{\rceil}

\newcommand{\regret}[3]{R_{#1}(#2,#3)}
\newcommand{\Reg}{\mathrm{Regret}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Surrogate Regret Bounds for Polyhedral Losses}
%Rate Transfers with Polyhedral Surrogate Losses
% \And separates authors; \AND separates with forced line break
\author{%
  Rafael Frongillo \\
  U. Colorado, Boulder \\
  \texttt{raf@colorado.edu} \\
  \And
  Bo Waggoner \\
  U. Colorado, Boulder \\
  \texttt{bwag@colorado.edu}
}

\begin{document}

\maketitle

\begin{abstract}
  Surrogate risk minimization is an ubiquitous paradigm in supervised machine learning, wherein a target problem is solved by minimizing a surrogate loss on a dataset.
  Surrogate regret bounds are a common tool to prove generalization rates for surrogate risk minimization.
  While surrogate regret bounds have been developed for certain classes of loss functions, such as proper losses, general results are relatively sparse.
  We provide two general results.
  The first gives a linear surrogate regret bound for any polyhedral (piecewise-linear convex) surrogate, meaning that surrogate generalization rates translate directly to target rates.
  The second shows that for sufficiently non-polyhedral surrogates, the regret bound is a square root, meaning fast surrogate generalization rates translate to slow rates for the target.
  Together, these results suggest polyhedral surrogates are weakly optimal in many cases.
  \raft{need to add keyword and/or mention: ``excess risk bounds'' (more common in the literature actually)}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
In supervised learning, our goal is to learn a hypothesis that solves some target problem given labeled data.
The problem is often specified by a target loss $\ell$ that is discrete in the sense that the prediction lies in a finite set, as with classification, ranking, and structured prediction.
As minimizing the target $\ell$ over a data set is typically NP-hard, we instead turn to the general paradigm of \emph{surrogate risk minimization}.
Here a surrogate loss function $L$ is chosen and minimized instead of the target $\ell$.
A link function $\psi$ then maps the surrogate predictions to target predictions.

Naturally, a central question in surrogate risk minimization is how to choose the surrogate $L$.
A minimal requirement is statistical \emph{consistency}, which roughly says that approaching the best possible surrogate loss will eventually lead to the best possible target loss.
A more precise goal is to understand the rate at which this minimization occurs, called the \emph{generalization rate}: how quickly the target loss approaches the best possible, in terms of the number of data points.
In other words, if we define the \emph{regret} of a hypothesis to be its performance relative to the Bayes optimal hypothesis, we wish to know how quickly the $\ell$-regret approaches zero.

Surrogate regret bounds are a common tool to prove generalization rates for surrogate risk minimization, by bounding the target regret as a function of the surrogate regret.
This bound allows generalization rates for surrogate losses---of which there are many, often leveraging properties of $L$ like strong convexity and smoothness---to imply rates for the target problem.
%The idea is that we have ample results to quantify the rate at which surrogate regret approaches zero, and if we could bound the target regret in terms of the surrogate regret, we would arrive at an overall generalization rate.
Roughly speaking, a surrogate regret bound takes the form
\begin{equation}
  \label{eq:surrogate-regret-bound-informal}
  \forall \text{ surrogate hypotheses } h, \quad \Reg_\ell(\psi\circ h) \leq \zeta(\, \Reg_L(h) \,)~,\qquad
\end{equation}
where the \emph{regret transfer function} $\zeta : \reals_+ \to \reals_+$ controls how surrogate regret translates to target regret.
For example, if we have a sequence of surrogate hypotheses $h_n$ achieving a fast rate $\Reg_L(h_n) = O(1/n)$, and we have a regret transfer of $\zeta(\epsilon) = \sqrt{\epsilon}$, then we can only guarantee a slower target rate of $\Reg_\ell(\psi \circ h_n) = O(1/\sqrt{n})$.
The ideal regret transfer is \emph{linear}, i.e. $\zeta(\epsilon) = C \cdot \epsilon$ for some $C > 0$, which would have directly translated the fast-rate guarantee to the target.

%As the above example illustrates, understanding the behavior of $\zeta$ near $0$ is crucial to analyzing the overall generalization rate; were $\zeta$ linear near $0$, the overall rate would have been a fast $O(1/n)$.
\raft{Maybe a good place to cite the IEEE rates paper}
Much is known about surrogate regret bounds for particular problems, especially for binary classification and related problems like bipartite ranking; see \S~\ref{sec:related-work} for a discussion of the literature.
General results which span multiple target problems, however, are sparse.
%, often forcing authors to prove bounds for specific surrogate losses from scratch, e.g.~\cite{ramaswamy2015hierarchical,ramaswamy2018consistent}.
Furthermore, we still lack guidance as to which surrogate will lead to the best overall target generalization rate, as this rate depends on both the surrogate generalization rate and the regret transfer function.
For example, consider the choice between a \emph{polyhedral} (piecewise-linear and convex) surrogate, like hinge loss, or a smooth surrogate like exponential or squared loss.
The surrogate rate for polyhedral losses will likely be a slower $O(1/\sqrt{n})$, whereas one typically achieves a faster $O(1/n)$ rate with smooth surrogates---but what is the tradeoff when it comes to the regret transfer function $\zeta$, which ultimately decides the target generalization rate?
To answer these types of questions, we need more general results.

Our first result is a general surrogate regret bound for polyhedral surrogates.
% Examples include hinge loss, the abstain loss of \bo{cite ramaswamy}, \bo{and others?}
Perhaps suprisingly, we show that \emph{every} consistent surrogate in this class achieves a linear regret transfer, meaning their surrogate generalization rates translate directly to target rates, up to a constant factor.
  % \begin{theorem}[Linear bounds for polyhedral surrogates]
\begin{theorem}[Linear bound for polyhedral]
  \label{thm:main-upper}
  Let $L$ be any polyhedral surrogate and $\psi$ any link function such that $(L,\psi)$ is consistent for the target loss $\ell$.
  Then $L$ satisfies Equation \ref{eq:surrogate-regret-bound-informal} with a linear $\zeta$.
  %Then there is a linear regret transfer function $\zeta$.
  % Then $(L,\psi)$ guarantee a linear regret transfer function $\zeta$ to $\ell$.
\end{theorem}
The recent embedding framework of~\citet{finocchiaro2019embedding} shows how to construct consistent polyhedral surrogates for any discrete target loss; Theorem~\ref{thm:main-upper} additionally endows these surrogates with linear regret bounds.
In particular, this result immediately applies to the Weston--Watkins surrogate~\cite{wang2020weston}, Lov\'asz hinge~\cite{yu2018lovasz,finocchiaro2019embedding}, and other polyhedral surrogates in the literature currently lacking regret bounds.

We also give a partial converse: If the surrogate is sufficiently ``non-polyhedral'', i.e. smooth and locally strongly convex, then $\zeta$ cannot shrink faster than a square root.
For these losses, therefore, even fast surrogate generalization rates would likely translate to slow rates for the target.
%This result could be viewed as a partial converse to the first, providing evidence that some element of piecewise-linearity is necessary for optimal surrogate regret bounds.
\begin{theorem}[Lower bound for ``non-polyhedral'']
  \label{thm:main-lower}
  Let $L$ be a locally strongly convex surrogate with locally Lipschitz gradient.
  If $L$ satisfies Equation \ref{eq:surrogate-regret-bound-informal} for a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%  If $\zeta$ is a transfer function in a surrogate regret bound for $L$ to a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
\end{theorem}
% Thus, polyhedral losses are not only order-optimal in terms of regret transfer, but a significant element of piecewise-linearity seems \emph{necesseary} for optimal transfer.
While it may seem intuitive that $\zeta=\sqrt{\cdot}$ for strongly convex losses, it is well-known that losses like exponential loss for binary classification also have $\zeta=\sqrt{\cdot}$, even though they are neither strongly convex nor strongly smooth~\cite{bartlett2006convexity}.
Theorem~\ref{thm:main-lower} offers an explanation for this phenomenon, and together with Theorem~\ref{thm:main-upper} clarifies an intriguing trend in the literature: $\zeta$ is typically either linear or square-root.

Taken together, our results shed light on the dichotomy expressed above: slower polyhedral surrogate rates of $O(1/\sqrt{n})$ will translate to the target unchanged, while most smooth surrogates, even if achieving a fast surrogate generalization rate of $O(1/n)$, will likely yield the same target rate of $O(1/\sqrt{n})$.
Our results therefore suggest an inherent tradeoff fast surrogate rates and optimal regret transfer, a phenomenon also observed by~\citet{mahdavi2014binary} for binary classification.
In particular, our results suggest that polyhedral losses may be weakly optimal in many cases.
% This result may be at odds with the smoothness and strong convexity properties often used to show fast convergence rates for convex surrogates \bo{cite?}.
% We leave to other work the possibility of an inherent tradeoff between fast surrogate convergence and optimal regret transfer; this question is explored to an extent in \bo{IEEE paper}.

\raft{I didn't get much out of this paragraph (apart from the first sentence), whereas I thought the next one was really useful.  I think I'll combine with a shortened version of the first, like ``Theorem 1 in \S, Theorem 2 in \S.  We make Theorem 1 constructive in \S...''}
The proof of Theorem \ref{thm:main-upper}, presented in \S~\ref{sec:upper}, relies on techniques from \emph{property elicitation} and the structure of optimal prediction sets for polyhedral losses.
In particular, a polyhedral loss has some finite set of reports which always contains a minimizer~\cite{finocchiaro2019embedding}.\raft{FYI, you were using our embedding dimension paper; I fixed throughout}
% This is related to the embeddings framework of \citet{finocchiaro2020embedding}.
Theorem \ref{thm:main-lower}, presented in \S~\ref{sec:lower}, uses curvature of the expected loss to construct a counterexample where convergence of target regret is quadratically slower than that of surrogate regret.

Our proof of the upper bound, Theorem \ref{thm:main-upper}, is nonconstructive.
We make it constructive in Section \ref{sec:constant}.
Specifically, a polyhedral loss's regret transfer function can be bounded by $\zeta(t) \leq \alpha \cdot t$ where the constant is of the form $\alpha = \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$.
Each component of $\alpha$ is constructed as a function of its respective object $\ell$, $L$, and $\psi$.
A corollary of the derivation is that a polyhedral loss $L$ and link $\psi$ are consistent for $\ell$ if and only if they satisfy the weak indirect elicitation criterion and $\psi$ is \emph{$\epsilon$-separated}, a condition introduced in \citet{finocchiaro2019embedding}.

\subsection{Literature on surrogate regret bounds}
\label{sec:related-work}

\raf{Basic outline, to flesh out tomorrow!}
\begin{itemize}
\item It all started with binary classification (Zhang), where we know the most.  Bartlett gave a nice closed form expression for $\zeta$ for margin losses, and Reid+W for the mostly broader class of composite proper losses.  So we have tight $\zeta$ functions for nearly every consistent surrogate.
\item By extension, we know a lot about close variants of binary classification, like cost-sensitive binary classification, binary classification with an abstain option~\cite{bartlett2008classification}, and bipartite ranking.
\item The next step is multiclass classification, yet commonly-used surrogates are actually not consistent.  For example, the common practice of one-versus-all hinge loss is actually consistent for multiclass classification with an abstain option; \citet{ramaswamy2018consistent} give a linear surrogate regret bound for this polyhedral surrogate, as well as two others for the same problem.  Related: hierarchical classification, which also has a polyhedral surrogate and linear regret bound.  Our results would have greatly simplified the development of these regret bounds; we describe how to recover the main BEP bound in \S.
\item Finally, structured prediction be pumpin out surrogates.... general results... but usually quadratic losses.  Show lower bounds like our Thm 2 in those two papers.
\end{itemize}

~\citep{zhang2004statistical,bartlett2006convexity,reid2009surrogate,mahdavi2014binary}

\bo{not sure if you wanted to keep this. Would rather move things here than inline if they're not that important to the narrative of the intro.}
Surrogate regret bounds have been well studied in some settings.
It is known that $\zeta$ \raf{examples: logistic, exponential, and hinge}.
Some general bounds are known for classes of surrogates, e.g. \raf{refs}.
Lacking, however, is a general understanding of what $\zeta$ one can expect from a surrogates, especially for arbitrary target losses.

% 1. super general
% 2. lower bound could explain why you get sqrt or linear
%   - take exp: not strong convex or strong smooth, so why sqrt??
% 3. polyhedral pipeline

Some references by setting.  P = polyhedral.  Surrogate regret bound: L = linear, Q = quadratic (sqrt).  $\leq$ = gives a ``lower bound'' for the surrogate regret too.
\begin{itemize}
\item Binary classification.
  \begin{itemize}
  \item \citet{zhang2004statistical} -- early analysis, didn't use ``surrogate'' but had some quantitative bounds that implied consistency
  \item \citet{bartlett2006convexity} -- tight bounds for \textbf{all margin losses}
  \item \citet{reid2009surrogate} -- tight bounds for \textbf{all proper composite losses}
  \item \citet{mahdavi2014binary} -- consider the generalization rate of the surrogate together with the ``excess risk bound'' (surrogate regret bound) (\textbf{and} the optimization actually), both as a function of the smoothness parameter of the surrogate.  They don't seem to give lower bounds on the excess risk bound though, just a degrading upper bound on it.  Still, very close to our work (and more generally the ``full stack'' idea!) in spirit.
  \end{itemize}

\item Weighted / cost-sensitive variants
  \begin{itemize}
  \item \citet{scott2011surrogate} -- weird example-dependent problem
  \item \citet{scott2012calibrated} -- general cost-sensitive binary classification
  \end{itemize}

\item Bipartite ranking (which is basically a set of pairwise binary classifications)
  \begin{itemize}
  \item \citet{agarwal2014surrogate} -- builds on binary classification losses (Mark + Bob), introduces strongly proper
  \item \citet{menon2016bipartite} -- different sort of regret bound, same problem
  \end{itemize}

\item Abstain and hierarchical
  \begin{itemize}
  \item[P,L] \citet{ramaswamy2018consistent} -- Abstain paper; shows bounds for 3 different surrogates, all \textbf{linear}, each ``from scratch''.
  \item[Q] \citet{yuan2010classification} -- Smooth? losses and characterizations for abstain; \textbf{binary classification only}
  \item[P,L] \citet{bartlett2008classification} -- The OG abstain paper, with a polyhedral hinge-like loss.  \textbf{Linear} regret bound.
  \item[P,L] \citet{ramaswamy2015hierarchical} -- Nested OvA surrogate; linear bound
  \end{itemize}

\item Ordinal regression
  \begin{itemize}
  \item \citet{pedregosa2017consistency} -- Pretty fascinating paper actually, about ordinal regression, i.e., when the label is the ordinal rank of the outcome, and how to construct surrogates.  For a class of surrogates based on binary classification, which is also cool, they get similar excess risk bounds to \citet{bartlett2006convexity}.
  \end{itemize}

\item List ranking
  \begin{itemize}
  \item[Q] \citet[Theorem 10]{ravikumar2011ndcg}
  \end{itemize}

\item Structured prediction (covering lots of target problems simultaneously)
  \begin{itemize}
  \item[Q,$\leq$] \citet[Theorem 7]{osokin2017structured} -- Prove excess risk bounds for a broad family of structured prediction problems, using a specific quadratic surrogate.  Also give an upper bound on the calibration function, i.e. a lower bound in the sense that we prove (Theorem 8).
  \item[Q] \citet{ciliberto2016consistent} -- Calibration function for a similar quadratic surrogate, for a similarly general class of structed pred problems.
  \item[Q,$\leq$] \citet[Theorem 4.4]{nowak2019general} -- Even more general than the previous Bach paper \cite{osokin2017structured} (e.g. includes ordinal regression).  Quadratic risk bound for a proper loss for the mean from a strongly-convex potential/entropy.
    Also give a matching lower bound in our sense (Theorem 4.5).
  \item[Q] \citet[Proposition 5]{blondel2019structured} -- Tighter losses, still quadratic, still quadratic excess risk.
  \end{itemize}

\item Polyhedral surrogates without risk/regret bounds
  \begin{itemize}
  \item \citet{wang2020weston} -- Weston--Watkins: calibrated for ordered partition
  \item \citet{yu2018lovasz} -- Lov\'asz hinge: calibrated for an abstain-like loss (from our paper)
  \item Top-$k$ losses, like \citet{lapin2016loss} and any others from \citet{yang2018consistency}
  \end{itemize}

\item Fast rates
  \begin{itemize}
  \item \citet{koren2015fast} -- fast rates for exp-concave surrogates ($O(d/n)$ where $d$ is the dimension) .  Note here exp-concavex is in the \emph{parameter}.  So $F(\theta) = L(h_\theta(x),y)$ should be exp-concave in $\theta$.  For $\alpha>0$ we say $f$ is $\alpha$-exp-concave if $x \mapsto \exp(-\alpha f(x))$ is concave.
  \item \citet{grunwald2015fast} -- Tim et al paper than Anish referenced!  
  \end{itemize}

\item Misc / didn't read
  \begin{itemize}
  \item Excess Risk Bounds for Exponentially Concave Losses \url{https://arxiv.org/pdf/1401.4566.pdf}
  \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}

%% Bo: I thought we could skip this sentence.
%In this section, we introduce our notation while recalling the motivation behind surrogate risk minimization and surrogate regret bounds.

\paragraph{Target problems.}
We consider supervised machine learning problems where data points have features in $\X$ and labels in $\Y$, a finite set.
The \emph{target problem} is to learn a hypothesis mapping $\X$ to a \emph{report} (or \emph{prediction}) \emph{space} $\R$, possibly different from $\Y$.\footnote{For example, $\R$ may be the set of rankings (permutations) of $\Y$.}
In this paper, we assume $\R$ is a finite set.
%For example, standard binary classification has $\Y = \R = \{-1,+1\}$, whereas in a ranking problem, $\R$ is the set of all permutations of $\Y$.
We suppose the target problem is specified by the \emph{target loss} $\ell:\R\to\reals^\Y_+$, which maps a report (prediction) $r \in \R$ to the vector $\ell(r)$ of nonnegative loss values for each label, i.e. $\ell(r) = (\ell(r)_y)_{y\in\Y}$.
Given a hypothesis $g: \X \to \R$ and a data point $(x,y) \in \X \times \Y$, the loss is $\ell(g(x))_y$.
We make a minimal non-redundancy assumption on $\ell$, formalized in Section \ref{subsec:elic}.
We may refer to $\ell$ as a \emph{discrete} target loss to emphasize our assumption that $\R$ and $\Y$ are finite.

\paragraph{Surrogate problems.}
We recall the surrogate risk minimization framework.
First, one constructs a \emph{surrogate loss function} $L:\reals^d\to\reals^\Y_+$.
Then, one minimizes it to learn a hypothesis of the form $h: \X \to \reals^d$.
Finally, one applies a \emph{link function} $\psi: \reals^d \to \R$.
If the learned hypothesis predicts $h(x) \in \reals^d$, then the implied prediction for the target problem is $\psi(h(x)) \in \R$.
For example, in binary classification, the target problem is given by $0$-$1$ loss, $\ell(r)_y = \ones\{r\neq y\}$, with surrogate losses typically defined over $\reals$.
Common surrogate losses include hinge loss and logistic loss.
These surrogates are typically associated with the link $\psi(u) = \mathrm{sign}(u)$, which maps negative predictions to $-1$ and nonnegative predictions to $+1$.

\paragraph{Surrogate regret bounds.}
The efficiency of a chosen surrogate is often quantified with a \emph{surrogate regret bound}.
%We define the \emph{regret} of a hypothesis $h: \X \to \reals^d$ as the expected loss relative to the best possible.
Formally, the \emph{surrogate regret} of $h$ with respect to data distribution $\D$ is given by $R_L(h;\D) = \E_{(X,Y)\sim\D} L(h(X))_Y - \inf_{h':\X\to\reals^d} \E_{(X,Y)\sim\D} L(h'(X))_Y$, where the minimum is taken over all measurable functions.
(One may equivalently assume the Bayes optimal hypothesis is in the function class.)
Given $h$, the \emph{target regret} of the implied hypothesis $\psi \circ h$ is $R_\ell(\psi\circ h;\D) = \E_{(X,Y)\sim\D} \ell(\psi(h(X)))_Y - \inf_{h':\X\to\R} \E_{(X,Y)\sim\D} \ell(h'(X))_Y$, where we recall that $\psi$ is the link function.

%% Addressed in intro instead
%A central concern for learning algorithms is the \emph{rate of convergence} of regret.
%If $h_n$ denotes an algorithm's output on $n$ data points, then a typical bound is of the form $R_L(h_n;\D) \leq O\left(\frac{1}{n^{\alpha}}\right)$, usually for some $\alpha$ between $\tfrac{1}{2}$ (``slow'') and $1$ (``fast'').
%However, such bounds for the surrogate leave open the question of the implied rate of convergence for the target problem.
%This is the question we address by studying \emph{surrogate regret bounds}.

A surrogate regret bound answers the following question: when we minimize the surrogate regret at a rate $f(n)$, given $n$ data points, how fast does the target regret diminish?
Formally, we say that $\zeta : \reals_+ \to \reals_+$ is a \emph{regret transfer function} if $\zeta$ is continuous at $0$ and satisfies $\zeta(0) = 0$.
Given a rerget transfer function $\zeta$, we say that $(L,\psi)$ guarantees a \emph{regret transfer of $\zeta$} for $\ell$ if
\begin{equation}
  \label{eq:surrogate-regret-bound}
  \forall \D, \forall h:\X\to\reals^d, \quad R_\ell(\psi\circ h;\D) \leq \zeta(\, R_L(h;\D) \,)~.
\end{equation}
We note that a classic minimal requirement for $(L,\psi)$ is that they be \emph{consistent} for $\ell$, which is the case if there exists any regret transfer $\zeta$ guaranteed by $(L,\psi)$ for $\ell$.\bo{cite}
In other words, consistency implies that $R_L(h;\D) \to 0 \implies R_{\ell}(\psi \circ h;\D) \to 0$, but says nothing about the relative rates.

%% Addressed in intro
%However, consistency alone could be compatible with slow transfer functions such as $\zeta(\epsilon) = \epsilon^{1/4}$.
%For such transfers, a ``fast'' learning rate of e.g. $O\left(\tfrac{1}{n}\right)$ in the surrogate problem would only transfer to a very slow rate of $O\left(\tfrac{1}{n^{1/4}}\right)$ in the target.
%On the other hand, we could hope to choose surrogates with as fast as a linear transfer rate, $\zeta(\epsilon) \leq C \cdot \epsilon$, implying $R_{\ell}(\psi \circ h;\D) \leq C \cdot R_L(h;\D)$.

%% Moved to intro
%Surrogate regret bounds have been well studied in some settings.
%It is known that $\zeta$ \raf{examples: logistic, exponential, and hinge}.
%Some general bounds are known for classes of surrogates, e.g. \raf{refs}.
%Lacking, however, is a general understanding of what $\zeta$ one can expect from a surrogates, especially for arbitrary target losses.

%% Addressed in intro
%We give two main results to deepen this understanding.
%The first is that polyhedral (piecewise-linear and convex) surrogates always yield \emph{linear} regret bounds.
%The second is that sufficiently convex and smooth losses (i.e., ``non-polyhedral'') require $\zeta(\epsilon) = \Omega(\sqrt{\epsilon})$.
%
%\begin{theorem}[Main surrogate regret bound]
%  Let $L$ be any polyhedral surrogate and $\psi$ any link function such that $(L,\psi)$ is consistent for the target loss $\ell$.
%  Then $(L,\psi)$ guarantee a linear regret transfer rate for $\ell$.
%  %Then there exists a surrogate regret bound with a linear transfer function $\zeta$.
%\end{theorem}
%
%\begin{theorem}[Main counterexample]
%  Let $L$ be a differentiable convex surrogate, with locally Lipschitz gradient, which is locally strongly convex.
%  If $L$ satisfies a regret transfer rate of $\zeta$ for a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%%  If $\zeta$ is a transfer function in a surrogate regret bound for $L$ to a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%\end{theorem}
%
%\raf{Paragraph on implication for rates here.  Maybe Huber example?}

\paragraph{Polyhedral surrogates.}
A function $f: \reals^d \to \reals$ is called \emph{polyhedral} if it can be written as a pointwise maximum of a finite set of affine functions~\citep{rockafellar1997convex}.\botodo{theorem/chapter?}
The epigraph of a polyhedral function is a polyhedral set, i.e. the intersection of a finite set of closed halfspaces; thus polyhedral functions are always convex.
We say a surrogate loss $L: \reals^d \to \reals_+^{\Y}$ is \emph{polyhedral} if, for each fixed $y \in \Y$, the loss as a function of prediction, i.e. $r \mapsto L(r)_y$, is polyhedral.

\subsection{Property elicitation and the conditional approach} \label{subsec:elic}
Our technical approach is based on property elicitation, which studies the structure of $\ell$ and $L$ as they relate to $\simplex$.
This involves a \emph{conditional} perspective.
Given a data distribution $\D$ on $\X \times \Y$, for $(X,Y) \sim \D$, we consider the format of possible target predictions $g(x) = r \in \R$; surrogate predictions $h(x) = u \in \reals^d$; and conditional distributions $\Pr[ \cdot \mid X] = p \in \simplex$.
By dropping the role of the hypotheses and the feature space $\X$, we can focus on the relationships of the functions $\ell(r)_y$ and $L(u)_y$ and the conditional distribution $p$.

In particular, the expected $L$-loss of prediction $u$ on distribution $p$ is simply the inner product $\inprod{p}{L(u)}$.
The \emph{Bayes risk} of $L$ is $\risk{L}(p) = \inf_{u \in \reals^d} \inprod{p}{L(u)}$, and similarly the Bayes risk of $\ell$ is $\risk{\ell}(p) = \inf_{r \in \R} \inprod{p}{\ell(r)}$.
We abuse notation to define the \emph{conditional surrogate regret} $R_L(u,p) = \inprod{p}{L(u)} - \risk{L}(p)$, and similarly the \emph{conditional target regret} $R_{\ell}(r,p) = \inprod{p}{\ell(r)} - \risk{\ell}(p)$.
Given a regret transfer function $\zeta$, we say $(L,\psi)$ guarantees a \emph{conditional regret transfer} of $\zeta$ for $\ell$ if, for all $p \in \simplex$ and all $u \in \reals^d$, $R_{\ell}(\psi(u),p) \leq \zeta(R_L(u,p))$.

The following is a direct result of Jensen's inequality, using that $R_L(h;\D) = \E_X R_L(h(X),p_X)$ where $p_X$ is the conditional distribution on $Y$ given $X$.
\begin{observation} \label{obs:transfer}
  If $(L,\psi)$ guarantee a conditional regret transfer of $\zeta$ for $\ell$, and $\zeta$ is concave, then $(L,\psi)$ guarantee a regret transfer of $\zeta$ for $\ell$.
\end{observation}

\paragraph{Property elicitation.}
To begin, we define a \emph{property}, which is essentially a statistic or summary of a distribution, such as the mode.
We then define what it means for a loss function to \emph{elicit} a property.
We write $\Gamma: \simplex \toto R$ as shorthand for $\Gamma: \simplex \to 2^{R} \setminus \{\emptyset\}$.

\begin{definition}[Property, level set]\label{def:property}
  A \emph{property} is a function $\Gamma:\simplex\toto R$, for some set $R$.
  The \emph{level set} of $\Gamma$ for report $r$ is the set $\Gamma_r = \{p \in \simplex : r \in \Gamma(p)\}$.
\end{definition}
The following definition applies both to target losses and surrogate losses.
\begin{definition}[Elicits, report space]
  \label{def:elicits}
  A loss function $f:R\to\reals^\Y_+$, where the domain $R$ is called the \emph{report space}, \emph{elicits} a property $\Gamma:\simplex \toto \R$ if
  \begin{equation}
    \forall p\in\simplex,\;\;\;\Gamma(p) = \argmin_{r \in R} \inprod{p}{f(r)}~.
  \end{equation}
\end{definition}
As each loss $f$ elicits a unique property, we refer to it as $\prop{f}$.
In particular, we will consistently use the notation $\Gamma = \prop{L}$ where $L$ is a surrogate loss with report space $\reals^d$, and we will use $\gamma = \prop{\ell}$ where $\ell$ is a target loss with finite report space $\R$.

We will assume throughout that the target loss is \emph{non-redundant}, meaning that all $r \in \R$ are possibly uniquely optimal.
Formally, we suppose that for all $r \in \R$, there exists $p \in \simplex$ with $\gamma(p) = \{r\}$.

\paragraph{Calibration and consistency.}
It is known that consistency implies the weaker condition of \emph{calibration}, which implies the still-weaker, but very useful, condition of \emph{indirect elicitation}.
To define these conditions, consider $L,\psi,\ell$ and $\gamma = \prop{\ell}$.
Recall that $\gamma(p)$ is the set of target reports that are optimal for $p$.
If $\psi(u) \not\in \gamma(p)$, then $u$ is a ``bad'' (suboptimal) surrogate prediction for $p$.
Let $B_{L,\psi,\ell}(p) = \{R_L(u,p) ~:~ \psi(u) \not\in \gamma(p)\}$, the set of expected surrogate losses resulting from bad predictions.
\begin{definition}[Indirectly elicits, Calibrated]
  A surrogate-link pair $(L,\psi)$ \emph{indirectly elicits} a target $\ell$ if $0 \not\in B_{L,\psi,\ell}(p)$ for all $p \in \simplex$.
  Meanwhile, $(L,\psi)$ is \emph{calibrated} for $\ell$ if $0 < \inf B_{L,\psi,\ell}(p)$ for all $p \in \simplex$.
\end{definition}
Equivalently, $(L,\psi)$ indirectly elicits $\ell$ if $u \in \Gamma(p) \implies \psi(u) \in \gamma(p)$.
Calibration is the slightly more subtle condition that there is a nonzero gap between the expected loss of a ``good'' report and that of any bad one.
\begin{fact}[\cite{bartlett2006convexity,steinwart2008support}]
  \label{fact:consistent-calibrated-elicits}
  Let $\ell: \R \to \reals_+^{\Y}$ be a target loss with finite $\R$ and $\Y$.
  If a surrogate and link $(L,\psi)$ are consistent for $\ell$, then $(L,\psi)$ are calibrated for $L$.
  If $(L,\psi)$ are calibrated for $\ell$, then they indirectly elicit $\ell$.
\end{fact}

%% move to inline where used
%\begin{definition}[Embedding a loss]\label{def:loss-embed}
%  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ if there exists some injective embedding $\varphi:\R\to\reals^d$ such that
%  (i) for all $r\in\R$ we have $L(\varphi(r)) = \ell(r)$, and (ii) for all $p\in\simplex,r\in\R$ we have
%  \begin{equation}\label{eq:embed-loss}
%    r \in \prop{\ell}(p) \iff \varphi(r) \in \prop{L}(p)~.
%  \end{equation}
%\end{definition}

\section{Upper Bound: Polyhedral Surrogates} \label{sec:upper}
In this section, we describe the proof of Theorem \ref{thm:main-upper}, that polyhedral surrogates guarantee linear regret transfers.
At a high level, the proof (whose details appear in Appendix \ref{app:upper}) has two parts:
\begin{enumerate}
  \item Fix $p \in \simplex$.
        Prove that there is some constant $\alpha_p$ such that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for all $u\in\reals^d$.
      \item Apply this result to each $p$ in a carefully chosen finite subset of $\simplex$.
        Argue that the maximum $\alpha_p$ from this finite set suffices.
\end{enumerate}
The first part will actually hold for any calibrated surrogate.
The second part relies on a crucial fact about polyhedral losses $L$: their properties $\Gamma=\prop{L}$ cover the whole simplex with only a finite number of polyhedral level sets.
In other words, although the prediction space $\reals^d$ is infinite, one can restrict to just a finite set of prediction points and always have a global minimizer of expected loss.

\paragraph{Part 1: Linear transfer for fixed $p$.}
First, we establish a linear transfer for any fixed conditional distribution $p$.
This holds for any calibrated surrogate, not necessarily polyhedral.
\begin{lemma} \label{lemma:fixed-p}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss and suppose the surrogate $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ are calibrated for $\ell$.
  Then for any $p \in \simplex$, there exists $\alpha_p \geq 0$ such that, for all $u \in \reals^d$,
    \[ R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p) . \]
\end{lemma}
While this result is nice, it is also somewhat nonconstructive and potentially loose.
In Section \ref{sec:constant}, we investigate the constant $\alpha_p$ for polyhedral losses in particular.

\paragraph{Part 2: Linear overall transfer function.}
Given Lemma \ref{lemma:fixed-p}, we might hope to obtain that, for any calibrated surrogate, there exists $\alpha = \sup_p \alpha_p$ such that $R_{\ell}(\psi(u),p') \leq \alpha R_L(u,p')$ for all $p'$.
However, we know this is false in general: not all surrogates yield linear regret transfers!
Indeed, for many surrogates, the supremum will be $+\infty$.

However, polyhedral surrogates have a special structure that allows us to show $\alpha$ is finite.
We first present some tools that hold for general surrogates, then the key implication from $L$ being polyhedral.

\begin{lemma} \label{lemma:refines}
  If $(L,\psi)$ indirectly elicits $\ell$, then $\Gamma = \prop{L}$ \emph{refines} $\gamma = \prop{\ell}$ in the sense that, for all $u \in \reals^d$, there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$.
\end{lemma}

The next lemma states that surrogate regret is linear in $p$ on any fixed level set $\Gamma_{u^*}$ of $\Gamma = \prop{L}$.
By combining this fact with Lemma \ref{lemma:refines}, we obtain that target regret is linear on these level sets as well.
\begin{lemma} \label{lemma:linear-on-levelset}
  Suppose $(L,\psi)$ indirectly elicits $\ell$ and let $\Gamma = \prop{L}$.
  Then for any fixed $u,u^* \in \reals^d$ and $r \in \R$, the functions $R_L(u,\cdot)$ and $R_{\ell}(r,\cdot)$ are linear in their second arguments on $\Gamma_{u^*}$.
\end{lemma}

We now arrive at the key property of polyhedral losses.
\begin{lemma} \label{lemma:polyhedral-finite}
  If $L: \reals^d \to \reals_+^{\Y}$ is polyhedral, then $\Gamma = \prop{L}$ has a finite set of level sets that union to $\simplex$. Also, these level sets are polytopes.
\end{lemma}

We are now ready to restate and sketch a proof of the main upper bound.
\begin{theorem}[Theorem \ref{thm:main-upper} restated] \label{thm:main-upper-details}
  Suppose the surrogate loss $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ are consistent for the target loss $\ell: \R \to \reals_+^{\Y}$.
  If $L$ is polyhedral, then $(L,\psi)$ guarantee a linear regret transfer for $\ell$, i.e. there exists $\alpha \geq 0$ such that, for all $\D$ and all measurable $h: \X \to \reals^d$,
    \[ R_{\ell}(\psi \circ h ; \D) \leq \alpha R_L(h ; \D) . \]
\end{theorem}
The key point of the proof is that, by Lemma \ref{lemma:polyhedral-finite}, the polyhedral loss $L$ has a finite set $U \subset \reals^d$ of predictions such that (a) for each $u \in U$, the level set $\Gamma_u$ is a polytope, and (b) $\cup_{u \in U} \Gamma_u = \simplex$.
Let $\mathcal{Q}$ be the finite set of all vertices of all these level sets.
For any vertex $q \in \mathcal{Q}$, we have a linear regret transfer for this fixed $q$ with constant $\alpha_q$ from Lemma \ref{lemma:fixed-p}.
Lemma \ref{lemma:linear-on-levelset} allows us to write the regret of a general $p$ as a convex combination of the regrets of its containing level set's vertices.
So we obtain a bound of $\alpha = \max_{q \in \mathcal{Q}} \alpha_q$.

\section{Lower bound} \label{sec:lower}

In this section, we show that if a surrogate loss is sufficiently ``non-polyhedral'', then the best regret transfer it can achieve is $\zeta(\epsilon)$ on the order of $\sqrt{\epsilon}$.
(Proofs are deferred to Appendix \ref{app:lower}.)
Specifically, we consider a relaxation of $L$ being both strongly smooth and strongly convex.

We next formalize this relaxation.
It will assume $L$ is strongly smooth everywhere and strongly convex around some \emph{boundary prediction for $L,\ell$}, which we define to be a $u_0 \in \reals^d$ such that, for some $r,r' \in \R$, there exists a distribution $p_0$ such that $p_0 \in \Gamma_{u_0} \cap \gamma_r \cap \gamma_{r'}$.
Here $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
The upshot of this definition is that for the conditional distribution $p_0$, $u_0$ minimizes expected $L$-loss, and it may link to either $r$ or $r'$, as either of these minimize expected $\ell$-loss.
We also note that an \emph{open neighborhood} of $u_0$ is an open set in $\reals^d$ containing $u_0$.

\begin{assumption} \label{assumption:lower}
  $L: \reals^d \to \reals^{\Y}_+$ is a surrogate and $\psi: \reals^d \to \R$ a link, consistent with the discrete target $\ell: \R \to \reals^{\Y}_+$, satisfying the following: (i) For all $y \in \Y$, the function $L(\cdot)_y$ is differentiable with a locally Lipschitz gradient.\bo{footnote def of locally lipschitz} (ii) For some $\alpha > 0$ and some $u_0 \in \reals^d$, a boundary prediction for $L,\ell$, the expected loss function $u \mapsto \inprod{p}{L(u)}$ is $\alpha$-strongly convex on an open neighborhood of $u_0$.
\end{assumption}

We observe that boundary points are guaranteed to exist if $(L,\psi)$ indirectly elicit $\ell$.
This holds because $\ell$ has a finite report space $\R$, so for example the fact that its full-dimensional level sets union to the simplex \bo{cite?} implies the existence of $r,r' \in \R$ and $p_0 \in \gamma_r \cap \gamma_{r'}$; and some surrogate prediction $u_0$ minimizes expected surrogate loss on $p_0$.

The following result is our main lower bound.
\begin{theorem}[Theorem \ref{thm:main-lower} restated.] \label{thm:main-lower-details}
  Suppose the surrogate loss $L$ and link $\psi$ satisfy a regret transfer of $\zeta$ for a target loss $\ell$.
  If $L$, $\psi$, and $\ell$ satisfy Assumption \ref{assumption:lower}, then there exists $c > 0$ such that, for arbitrarily small $\epsilon$, $\zeta(\epsilon) \geq c \sqrt{\epsilon}$.
\end{theorem}
By ``arbitrarily small'', we mean that for any $\epsilon_0 > 0$, we can find $\epsilon$ with $0 < \epsilon < \epsilon_0$ such that $\zeta(\epsilon) \geq c \sqrt{\epsilon}$.
Thus, there is no neighborhood of zero where $\zeta(\epsilon)$ shrinks faster than on the order of $\sqrt{\epsilon}$.

The key idea of the proof is to fix a boundary report $u_0$ and consider a sequence of distributions $p_{\lambda} \to p_0$ as $\lambda \to 0$.
Target regret $R_{\ell}(\psi(u_0),p_{\lambda})$ shrinks linearly in $\lambda$, but surrogate regret $R_L(u_0,p_{\lambda})$ shrinks quadratically.
This holds roughly because expected surrogate loss is strongly convex in some neighborhood of $u_0$ and is strongly smooth elsewhere.\footnote{We note that a tempting approach instead is to fix a distribution $p_0$ and consider a sequence of predictions. But, we know from Lemma \ref{lemma:fixed-p} that this will fail: for any consistent surrogate and any fixed $p_0$, there is a linear regret transfer if we restrict to $p_0$.}

We observe that Assumption \ref{assumption:lower} captures a wide range of surrogate losses beyond those that are strongly convex and strongly smooth.
For example, exponential loss is not strongly convex, but does satisfy Assumption \ref{assumption:lower}. \bo{more to say?}

\section{Unpacking the constant for polyhedral losses} \label{sec:constant}

Our main upper bound for polyhedral losses, Theorem \ref{thm:main-upper}, is mostly nonconstructive: it says polyhedral surrogates have some $\alpha$ such that $R_{\ell}(\psi \circ h; \D) \leq \alpha \cdot R_L(h;\D)$.
In this section, we make the result constructive by unpacking the implications of consistency for polyhedral surrogates.
Proofs appear in Appendix \ref{app:constant}.

We will give a bound on the constant of the form $\alpha \leq \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$, i.e. one component each relating to the structure of the target loss, surrogate loss, and link.
First, we will derive $\alpha_L$ using the theory of \emph{Hoffman constants} from linear optimization.
Then, we will show that \emph{any} link for a consistent polyhedral loss must satisfy a condition called $\epsilon$-separation, introduced in \cite{finocchiaro2019embedding}.
We will set $\alpha_{\psi} = \frac{1}{\epsilon}$.
With $\alpha_{\ell}$ simply an upper bound on the target loss, we will finally prove $\alpha \leq \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$.

\paragraph{Hoffman constants.}
The intuition we explore here is the linear growth rate of a given polyhedral function as we move away from its optimal set.
In particular, consider the expected loss $u \mapsto \inprod{p}{L(u)}$.
Expected loss should grow linearly with distance from the optimal set, which happens to be $\Gamma(p)$.
In particular, the worst growth rate is governed roughly by the smallest slope of the function in any direction.
This is captured by the \emph{Hoffman constant} of an associated system of linear inequalities.
In Appendix \ref{app:constant}, we use Hoffman constants to obtain the following result, where we define $H_{L,p}$.
\begin{lemma} \label{lemma:hoffman-polyhedral}
  Let $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral loss with $\Gamma = \prop{L}$.
  Then for any fixed $p$, there exists some smallest constant $H_{L,p} \geq 0$ such that $d_{\infty}(u,\Gamma(p)) \leq H_{L,p} R_L(u,p)$ for all $u \in \reals^d$.
\end{lemma}

To foreshadow the usefulness of $H_{L,p}$, recall the proof strategy of the upper bound: use Lemma \ref{lemma:fixed-p} to obtain a constant $\alpha_p$ for each $p$; then carefully choose a finite set of $p$'s and take the maximum constant.
In the same way, we will be able to set the ``loss'' constant $\alpha_L = \max_p H_{L,p}$.
First, however, we will need to investigate separated links.

\paragraph{Separated link functions.}
A link $\psi$ is $\epsilon$-separated if, when $\psi(u^*)$ is an optimal target prediction and $\psi(u)$ is not, then $\|u^* - u\|_{\infty} > \epsilon$.
The definition was introduced in \cite{finocchiaro2019embedding} and is rephrased in terms of the properties $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.

\begin{definition}[Separated Link]\label{def:sep-link}
  Let properties $\Gamma:\simplex\toto\reals^d$ and $\gamma:\simplex\toto\R$ be given.
  We say a link $\psi:\reals^d\to\R$
  % Let $L:\reals^d \to \reals^\Y_+$, $\gamma:\simplex\toto\R$, and link $\psi:\reals^d\to\R$ be given.
  is \emph{$\epsilon$-separated} if for all $u\in\reals^d$ with $\psi(u)\notin\gamma(p)$, we have $d_\infty(u,\Gamma(p)) > \epsilon$, where $d_\infty(u,A) \defeq \inf_{a\in A} \|u-a\|_\infty$.
\end{definition}

Recall from the previous subsection that Hoffman constants allowed us to show that polyhedral expected losses grow linearly as we move away from the optimal set.
Now, $\epsilon$-separated link functions allow us to guarantee that we move at least a constant distance from the optimal set before we start linking to an ``incorrect'' report $r' \in \R$.
But when does a polyhedral loss have an associated $\epsilon$-separated link?

\cite{finocchiaro2019embedding} showed that in fact, if $L$ indirectly elicits $\ell$, then there always exists an $\epsilon$-separated link $\psi$ such that $(L,\psi)$ is consistent for $L$.
Here, we strengthen this result much further, and give a converse that we believe is of independent interest: if $L$ indirectly elicits $\ell$, then \emph{every} link making $L$ consistent is $\epsilon$-separated for some $\epsilon > 0$.
\raft{Randomly came across Lemma 6 of~\citet{tewari2007consistency}, which seems to prove a similar result.  To cite properly later...}

\begin{lemma}\label{lemma:calibrated-eps-sep}
  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$, discrete loss $\ell:\R\to\reals^\Y_+$, and link $\psi:\reals^d\to\R$ be given such that $(L,\psi)$ is calibrated with respect to $\ell$.
  Then there exists $\epsilon>0$ such that $\psi$ is $\epsilon$-separated with respect to   $\Gamma \defeq \prop{L}$ and $\gamma \defeq \prop{\ell}$.
\end{lemma}

\paragraph{Combining the loss and link.}
We can now follow the general proof strategy of the main upper bound, but constructively.
Given Lemma \ref{lemma:calibrated-eps-sep}, we can make the following definition.
\begin{definition}[$\epsilon_{\psi}$]
  If $(L,\psi)$ are calibrated for $\ell$, then let $\epsilon_{\psi} \defeq \sup \{ \epsilon ~:~ \text{$\psi$ is $\epsilon$-separated}\}$.
\end{definition}
We will also need an upper bound on the regret of $\ell$.
\begin{definition}[$C_{\ell}$]
  Given a discrete target loss $\ell: \R \to \reals_+^{\Y}$, define $C_{\ell} = \max_{r,r' \in \R, y \in \Y} \ell(r)_y - \ell(r')_y$.
\end{definition}
Observe that $C_{\ell} = \max_{r,p} R_{\ell}(r,p)$.
\bo{needs justification?}

For the proof, we first give a bound for any fixed $p$.
\begin{lemma} \label{lemma:separated-constant-p}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss, $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral surrogate loss, and $\psi: \reals^d \to \R$ a link function.
  If $(L,\psi)$ indirectly elicit $\ell$ and $\psi$ is $\epsilon$-separated, then for all $u$ and $p$,
    \[ R_{\ell}(\psi(u),p) \leq \frac{C_{\ell} H_{L,p}}{\epsilon} R_L(u,p) . \]
\end{lemma}

Recall that Lemma \ref{lemma:polyhedral-finite} gave that, if $L$ is polyhedral, then $\Gamma = \prop{L}$ has a finite set of full-dimensional level sets, each a polytope, that union to the simplex.
\begin{definition}[$H_L$]
  Given a polyhedral surrogate loss $L: \reals^d \to \reals_+^{\Y}$, let $\mathcal{Q}$ be the set of all vertices of the full-dimensional level sets of $\Gamma = \prop{L}$, and define $H_L \defeq \max_{p \in \mathcal{Q}} H_{L,p}$.
\end{definition}

\begin{theorem}[Constructive linear transfer] \label{thm:separated-constant}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss, $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral surrogate loss, and $\psi: \reals^d \to \R$ a link function.
  If $(L,\psi)$ are consistent for $\ell$, then
    \[ (\forall h,\D) \quad R_{\ell}(\psi \circ h ; \D) \leq \frac{C_{\ell} H_L}{\epsilon_{\psi}} R_L(h ; \D) ~. \]
\end{theorem}
The proof closely mirrors the proof of the nonconstructive upper bound, Theorem \ref{thm:main-upper}, but using the constructive constants $C_{\ell}, H_L, \epsilon_{\psi}$ derived above.
To summarize, we have shown the following.
\begin{theorem} \label{thm:tfae}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss, $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral surrogate loss, and $\psi: \reals^d \to \R$ a link function.
  The following are equivalent:
  \begin{enumerate}
    \item $(L,\psi)$ are consistent for $\ell$.
    \item $(L,\psi)$ indirectly elicit $\ell$ and $\psi$ is $\epsilon$-separated for some $\epsilon > 0$.
    \item $(L,\psi)$ guarantee a linear regret transfer for $\ell$.
  \end{enumerate}
  Furthermore, if any of the above hold, then in particular $(L,\psi)$ guarantee the regret transfer $\zeta(t) = \left(\frac{C_{\ell} H_L}{\epsilon_{\psi}}\right) t$.
\end{theorem}

\subsection{Further tightening}
While our goal is not to focus on exact constants for specific problems, we remark on a few ways Theorem \ref{thm:separated-constant} may be loose and how one could compute the tightest possible constant $\alpha^*$ for which $R_{\ell}(\psi \circ h;\D) \leq \alpha^* R_L(h;\D)$ for all $h$ and $\D$.
In general, for a fixed $p$, there is some smallest $\alpha_p^*$ such that $R_{\ell}(\psi(u),p) \leq \alpha_p^* R_L(u,p)$ for all $u$.
Then, it follows from our results that $\alpha^* = \max_{p \in \mathcal{Q}} \alpha_p^*$ for the finite set $\mathcal{Q}$ used in the proof, i.e. the vertices of the full-dimensional level sets of $\Gamma = \prop{L}$.

Above, we bounded $\alpha_p^* \leq \frac{C_{\ell} H_{L,p}}{\epsilon_{\psi}}$.
The intuition is that some $u$ at distance $\geq \epsilon_{\psi}$ from $\Gamma(p)$, the optimal set, may link to a ``bad'' report $r = \psi(u) \not\in \gamma(p)$.
The rate at which $L$ grows is at least $H_{L,p}$, so the surrogate loss at $u$ may be as small as $\frac{\epsilon_{\psi}}{H_{L,p}}$, while the target regret may be as high as $C_{\ell} = \max_{r',p'} R_{\ell}(r',p')$.
The ratio of regrets is therefore bounded by $\frac{H_{L,p} C_{\ell}}{\epsilon_{\psi}}$.

The tightest possible bound, on the other hand, is $\alpha^* = \sup_{u: \psi(u) \not\in \gamma(p)} \frac{R_{\ell}(\psi(u),p)}{R_L(u,p)}$.
This can be smaller if the values of numerator and denominator are correlated across $u$.
For example, $u$ may only be $\epsilon_{\psi}$-close to the optimal set when it links to reports $\psi(u)$ with lower target regret; or $L$ may have a smaller slope in the direction where the link's separation is larger than $\epsilon$.

To illustrate with a concrete example, consider the \emph{binary encoded predictions (BEP) surrogate} of \cite{ramaswamy2018consistent} for the abstain target loss, $\ell(r,y) = \frac{1}{2}$ if $r = \bot$, otherwise $\ell(r,y) = \ones[r \neq y]$.
The surrogate involves an injective map $B: \Y \to \{-1,1\}^d$ for $d = \lceil \log_2 |\Y| \rceil$.
It is $L(u)_y = \max_{j=1\dots d} (1 - u_d B(y)_d)_+$, where $(\cdot)_+$ indicates taking the maximum of the argument and zero.
The associated link is $\psi(u) = \bot$ if $\min_{j=1\dots d} |u_j| \leq \tfrac{1}{2}$, otherwise $\psi(u) = \argmin_{y \in \Y} \|B(y) - u\|_{\infty}$.

One can show that for $p = \delta_y$, i.e. the distribution with full support on some $y \in \Y$, $L(u)_y = d_{\infty}(u,\Gamma(p))$ exactly, giving $H_{L,p} = 1$.
It is almost immediate that $\epsilon_{\psi} = \tfrac{1}{2}$.
Meanwhile, $R_{\ell}(r,p) \leq 1$, giving us an upper bound $\alpha^*_p \leq \frac{(1)(1)}{1/2} = 2$.
In fact, this is slightly loose; the exact constant, given in \cite{ramaswamy2018consistent} is $1$.
The looseness stems from the fact that for $p = \delta_y$, the closest reports $u$ to the optimal set, i.e. at distance only $\epsilon_{\psi} = \tfrac{1}{2}$ away, do not link to reports maximizing target regret; they link to the abstain report $\bot$, which has regret only $\tfrac{1}{2}$.
With this correction, and an observation that all $u$ linking to reports $y' \neq y$ are at distance at least $\tfrac{3}{2}$ from $\Gamma(p)$, we restore the tight bound $\alpha^*_p \leq 1$.
A similar but slightly more involved calculation can be carried out for the other vertices $p \in \mathcal{Q}$, which turn out to be all vertices of the form $\tfrac{1}{2} \delta_y + \tfrac{1}{2} \delta_{y'}$.

\paragraph{Different norms.}
This paper uses $L_{\infty}$ norm for the minimum-slope $H_L$ and the $\epsilon$-separated link.
One reason is that it is easiest to prove the former by reduction to existence of Hoffman constants; another reason is compatibility of both with the abstain loss example above.
However, all definitions hold for other $L_p$ norms and so does the main upper bound, as existence of an $H_L$ and $\epsilon_{\psi}$ in $L_{\infty}$ imply existence of constants in $L_p$ for $p \geq 1$.
The optimal constant may change if the distance measure changes.

\section{Broader Impacts}
As a theoretical work, this papers' impacts are mostly in the form of downstream research and applications.
Instead of direct applications, we anticipate this work leading to more investigation of surrogate losses to improve discrete prediction tasks.
It may inform practitioners' choices of which surrogate losses they use for supervised learning tasks.
Of course, such machine learning tasks can be solved for ethical or unethical purposes.
We do not know of particular risk of negative impacts of this work beyond risks of supervised machine learning in general.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Acknowledgements (automatically hidden in submission)
\begin{ack}
  \raf{Stephen Becker for help with the strong convexity stuff}
  \raf{Nishant Mehta for many discussions}
  \raf{Jessie Finocchiaro for discussions}
\end{ack}

%\section*{References}

\bibliographystyle{plainnat}
\bibliography{diss,extra}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Checklist}

%%% BEGIN INSTRUCTIONS %%%
%The checklist follows the references.  Please
%read the checklist guidelines carefully for information on how to answer these
%questions.  For each question, change the default \answerTODO{} to \answerYes{},
%\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
%justification to your answer}, either by referencing the appropriate section of
%your paper or providing a brief inline description.  For example:
%\begin{itemize}
%  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
%  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
%  \item Did you include the license to the code and datasets? \answerNA{}
%\end{itemize}
%Please do not modify the questions and only use the provided macros for your
%answers.  Note that the Checklist section does not count towards the page
%limit.  In your paper, please delete this instructions block and only keep the
%Checklist section heading above along with the questions/answers below.
%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerYes{}
  \item Did you describe the limitations of your work?
    \answerYes{}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerYes{}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerYes{}
\end{enumerate}

\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerYes{}
	\item Did you include complete proofs of all theoretical results?
    \answerYes{}
\end{enumerate}

\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerNA{}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerNA{}
	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerNA{}
	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerNA{}
\end{enumerate}

\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerNA{}
  \item Did you mention the license of the assets?
    \answerNA{}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerNA{}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerNA{}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerNA{}
\end{enumerate}

\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerNA{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerNA{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerNA{}
\end{enumerate}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{appendix}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: