### 2021-10-16 rates-neurips-21/talk-recorded/main.tex
started rates talk

\documentclass[t,mathserif,10pt,aspectratio=1610]{beamer}
% option "t" means align top
% option "handout" for handout mode (each frame takes only one page)
% option "notes" to add notes pages

\setbeamertemplate{footline}[frame number]

\mode<presentation>
{
  \usetheme{bo}
}

\newcommand{\mycite}[1]{{\footnotesize [\textit{#1}]}}

\newcommand{\centergraphics}[2]{
  \begin{center}
  \includegraphics[width=#1\textwidth]{#2}
  \end{center}} % end newcommand

\newcommand{\centerboxgraphics}[2]{
  \begin{center}
  \fbox{\includegraphics[width=#1\textwidth]{#2}}
  \end{center}} % end newcommand

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\include{defs}

\usepackage{adjustbox}

\title{\Large\textbf{Surrogate Regret Bounds for Polyhedral Losses}}
%\titlegraphic{
%\includegraphics[width=.35\textheight]{figs/logo}
%}

\author[Bo and Raf]{{\textbf{Rafael Frongillo} and \textbf{Bo Waggoner}} \and {University of Colorado, Boulder}}
\date{\textbf{NeurIPS 2021}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The short version}{}
  TODO: switch red and blue
  Surrogate risk minimization:
  \pause
  \begin{itemize}[<+->]
    \item Goal: optimize \blueem{discrete ``target''} loss $\ell(r,y)$  \hintright{e.g. classification, structured predictions, \dots}  \\
          \hint{$r = $ prediction in a finite set, $y = $ label}
    \item Approach: optimize \redem{continuous ``surrogate''} loss $L(u,y)$  \hintright{then ``link'' to target prediction $r = \psi(u)$} \\
          \hint{$u = $ prediction in $\reals^d$}
  \end{itemize}

  \only<3>{
    \centergraphics{0.9}{figs/hinge}
  }

  \pause
  \vskip2em
  Question: when does (quick) \redem{surrogate} convergence imply (quick) \blueem{target} convergence?  \\
  \hint{``Surrogate regret bounds'' or ``regret transfer rates''}

  \pause
  \vskip2em
  This paper: \textbf{polyhedral surrogates}  \hintright{piecewise-linear and convex}
  \begin{enumerate}[<+->]
    \item \emph{All} polyhedral surrogates have \textbf{linear} regret transfer rates!
    \item \emph{All} sufficiently ``non-polyhedral'' surrogate transfers are \textbf{quadratically slower}.
  \end{enumerate}

  \pause
  \vskip2em
  \textbf{Polyhedral:} \redem{Surrogate regret} $\leq \epsilon$ $\implies$ \blueem{target regret} $\leq O(\epsilon)$  \\
  \pause \textbf{Non-polyhedral:} Sometimes, \redem{surrogate regret} $\leq \epsilon$ yet \blueem{target regret} $\geq \Omega(\sqrt{\epsilon})$.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}