### 2021-10-22 rates-neurips-21/appendix.tex

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "rates"
%%% End: