### 2021-12-03 rates-neurips-21/neurips-poster/main.tex
pass for wordiness

\documentclass[t,mathserif,12pt,aspectratio=1610,handout]{beamer}
% option "t" means align top
% option "handout" for handout mode (each frame takes only one page)
% option "notes" to add notes pages

\mode<presentation>
{
  \usetheme{bo-poster}
}
\newcommand{\enumitem}[1]{{\usebeamercolor[bg]{item projected}%
  \raisebox{1.5pt}{\colorbox{bg}{\color{fg}#1}}}}%

\newcommand{\mycite}[1]{{\footnotesize [\textit{#1}]}}

\newcommand{\centergraphics}[2]{
  \begin{center}
  \includegraphics[width=#1\textwidth]{#2}
  \end{center}} % end newcommand

\newcommand{\centerboxgraphics}[2]{
  \begin{center}
  \fbox{\includegraphics[width=#1\textwidth]{#2}}
  \end{center}} % end newcommand

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\include{defs}

\usepackage{adjustbox}

\title{\Large\textbf{Surrogate Regret Bounds for Polyhedral Losses}}
%\titlegraphic{
%\includegraphics[width=.35\textheight]{figs/logo}
%}

\author[Bo and Raf]{{\textbf{Rafael Frongillo} and \textbf{Bo Waggoner}} \and {University of Colorado, Boulder}}
\date{\textbf{NeurIPS 2021}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The short version}{}
  Surrogate risk minimization for supervised learning:
  \pause
  \begin{itemize}[<+->]
    \item Goal: optimize \redem{discrete ``target''} loss $\ell(r,y)$  \hintright{classification, structured prediction}  \\
          \hint{$r = $ prediction, $y = $ label}
    \item Approach: optimize \blueem{continuous ``surrogate''} loss $L(u,y)$  \hintright{then ``link'' to target prediction $r = \psi(u)$} \\
%          \hint{$u = $ prediction in $\reals^d$}
  \end{itemize}

  \vskip1em
  \textbf{Q:} When does (quick) \blueem{surrogate} convergence imply (quick) \redem{target} convergence?  \\
  \hint{``Surrogate regret bounds'' or ``regret transfer rates''}

  \only<7>{
    {\Large \[ \Reg_{\ell}(\psi \circ h) \leq \zeta \Big( \Reg_L(h) \Big) \] }
    \hint{Regret = ``excess risk'' over Bayes optimal}
  }

  \vskip1em
  This paper: \textbf{polyhedral surrogates}  \hintright{piecewise-linear and convex}
\end{frame}

\begin{frame}{~}{}
  \vskip-1em
  \centerline{\includegraphics[width=0.65\linewidth]{figs/hinge}}
  \vskip1em

  \enumitem{\bf Theorem 1:} All polyhedral surrogates have \textbf{linear} regret transfer rates!

  \smallskip
  \textbf{Polyhedral:} \blueem{surrogate regret} $\leq \epsilon$ $\implies$ \redem{target regret} $\leq O(\epsilon)$.

  \vskip1em

  \enumitem{\bf Theorem 2:} Sufficiently ``non-polyhedral'' transfers are \textbf{quadratically slower}.

  \smallskip
  \textbf{Non-polyhedral:} Can have \blueem{surrogate regret} $\leq \epsilon$ yet \redem{target regret} $\geq \Omega(\sqrt{\epsilon})$.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Background: surrogate risk, polyhedral losses}{}
  \textbf{Data:} $(x,y) \in \X \times \Y$   \hintright{from distributions $\D$}

  \vskip1em
  \pause
  \textbf{Hypotheses:} $g: \X \to \R$   \hintright{example soon}

  \vskip1em
  \pause
  \redem{Discrete target loss:} $\ell: \R \times \Y \to \reals_{\geq 0}$.  \hintright{discrete: $\R$ is finite}  \\
  \hint{Classification ($0-1$ loss), ranking, top-$k$}

  \pause
  \pause
  \vskip1em
  \textbf{Regret:} $\Reg_{\ell}(g;\D) := \E_{x,y \sim \D} \left[ \ell(g(x), y) - \ell(g^*(x), y) \right]$.  \hintright{$g^* = $ Bayes optimal}

  \pause
  \vskip1em
  \blueem{Continuous surrogate loss:} $L: \reals^d \times \Y \to \reals_{\geq 0}$.  \hintright{for some $d$} \\
  \hint{Hinge loss for classification, BEP surrogate (Ramaswamy et al. 2018)}

  \vskip1em
  \blueem{Polyhedral surrogate} $L: \reals^d \times \Y \to \reals_{\geq 0}$: \\
  pointwise maximum of a finite set of affine functions.

  \pause
  \vskip1em
  Recent work: \blueem{polyhedral surrogates} are natural convexifications of \redem{discrete losses}.
  \begin{itemize}[<+->]
    \item Finocchiaro, Frongillo, Waggoner 2019, 2020; applications e.g. Wang, Scott 2020.
%    \begin{itemize}
%      \item Every \redem{discrete loss} $\ell$ has a consistent \blueem{polyhedral surrogate} $(L,\psi)$.
%      \item Further, that surrogate is obtained by ``embedding'' $\ell$ in $\reals^d$.
%    \end{itemize}
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{~}{}
  Desirable surrogates:
  \begin{itemize}
    \item Optimizable.  \hintright{Convex, strongly convex, etc.}
    \item Generalization bound, convergence rate.
%  \textbf{2.} Generalization bound; convergence rate. \\
%  \hint{e.g. with $n$ data points, $\Reg_L(h;\D) \leq O(\sqrt{1/n})$.}
    \item \textbf{Connection to \redem{target}...}
  \end{itemize}

  \vskip1em
  \textbf{Regret transfer function} (\emph{calibration function}) $\zeta: \reals_{\geq 0} \to \reals_{\geq 0}$:  \hintright{continuous at zero, $\zeta(0) = 0$}

  \[ \Reg_{\ell}(\psi \circ h ; \D) ~ \leq ~ \zeta\Big( \Reg_L(h ; \D) \Big) . \]

  \pause
  \vskip1em
  If \textbf{any such $\zeta$} exists, $(L,\psi)$ is \textbf{consistent} for $\ell$.  \\
  \hint{As $\Reg_L(h^{(n)}; \D) \to 0$, we have $\Reg_{\ell}^{(n)}(\psi \circ h^{(n)} ; \D) \to 0$.}

  \vskip1em
  \pause
  Ideally: \textbf{fast} convergence, e.g. $\zeta(\epsilon) = O(\sqrt{\epsilon})$ (good), ~~ $O(\epsilon^{3/4})$ (better), ~~ $O(\epsilon)$ (best).

  \vskip1em
  \pause
  Prior work: binary classification (Zhang et al. 2004, etc.); bipartite ranking (Agarwal 2014, etc.); multiclass classification (Duchi et al. 2018, etc.); hierarchical classification (Ramaswamy et al. 2015); strongly convex surrogates (Nowak-Vila et al. 2019, etc.).
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Results}{}
  \begin{block}{Theorem 1}
    Suppose $L$ is polyhedral and $(L,\psi)$ are consistent for the discrete target $\ell$.
    Then there exists $C > 0$ such that, for all $\D$ and $h$,
    \[ \Reg_{\ell}(\psi \circ h; \D) \leq C \cdot \Reg_L(h; \D) . \]
  \end{block}

  \pause
  \vskip1em
  \emph{Proof sketch.}
  Fix $x$; let $p,q \in \simplex$ be distributions of $y$ given $x$.
  \begin{itemize}[<+->]
    \item Suffices to prove that for all $p$ and $u$, $\Reg_{\ell}(\psi(u);p) \leq C \cdot \Reg_L(u;p)$.
    \item $\forall q$, $\exists \alpha_q > 0$ such that $\Reg_{\ell}(\psi(u); q) \leq \alpha_q \cdot \Reg_L(u; q)$.
    \item Polyhedral losses have a \textbf{finite structure}: \hintright{related to embeddings framework}
    \begin{itemize}
      \item There is a finite $U \subseteq \reals^d$ such that, for all $p$, $U$ contains an optimal prediction.
      \item These $U$ partition $\simplex$ into finitely many polytopes.
    \end{itemize}
    \item Regret is linear on each polytope.
    \item Therefore $\Reg_L(u ; p)$ is a convex combination over the corners.
    \item So we can take $C = \max_{\text{corners $q$}} \alpha_q$.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{~}{}
  \begin{block}{Theorem 2}
    Let $L$ be a locally strongly convex surrogate with locally Lipschitz gradient.
    Suppose $(L,\psi)$ is consistent for $\ell$ with:
    \[  \Reg_{\ell}(\psi \circ h) \leq \zeta\Big( \Reg_L(h) \Big) . \]
   Then there exists $c > 0$ such that, for all small enough $\epsilon > 0$,
    \[ \zeta(\epsilon) \geq c \sqrt{\epsilon} . \]
  \end{block}

  \pause
  \emph{Proof idea:}
  \begin{itemize}
    \item Fix a ``boundary'' prediction $u_0$.
    \item Consider a conditional distributions $\{p_{\lambda} : 0 \leq \lambda \leq 1\}$
    \item \redem{Target regret} shrinks linearly as $\lambda \to 0$, but \blueem{surrogate regret} shrinks with $\sqrt{\lambda}$.
  \end{itemize}

  \pause
  Examples: exponential loss, Huber loss \hintright{via strengthenings}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Investigating the constant in Theorem 1}{}
  $C = \max_{\text{corners $q$}} \alpha_q$.

  \pause
  \vskip1em
  What is $C$?

  \pause
  \vskip1em
  Can bound $C \leq \beta_{L} \cdot \beta_{\ell} \cdot \beta_{\psi}$.

  \pause
  \vskip1em
  \begin{itemize}
    \item $\beta_L$ comes from \textbf{Hoffman constants}.  \hintright{minimum slope of polyhedral losses}
    \item $\beta_{\ell}$ comes from a simple maximum possible regret.
    \item $\beta_{\psi} = \frac{1}{\epsilon}$ where the link is $\epsilon$-separated. \hintright{see embedding framework}
  \end{itemize}
\end{frame}

\begin{frame}{Big picture}{}
  \blueem{Polyhedral surrogate losses} are nice:
  \begin{itemize}
    \item Consistent polyhedral surrogates always exist
    \item Embedding framework
    \item \textbf{(This work) Always satisfy linear regret transfer}  \hintright{a.k.a. calibration functions}
  \end{itemize}

  \pause
  \vskip2em
  Next questions:
  \begin{itemize}
    \item Are they good for the whole pipeline? (optimization + generalization)
    \item Applications...
    \item e.g. low-dimensional or otherwise ``nice'' polyhedral surrogates
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}