### 2022-05-10 poly-embed-journal/tikz/lovasz-links.tex
Small figure touchups and writing edits in 5.1

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsfonts}
\usetikzlibrary{calc}
\newcommand{\R}{\mathcal{R}}

\begin{document}

\definecolor{deepblue}{HTML}{335C67}
\definecolor{champagne}{HTML}{FFF3B0}
\definecolor{softyellow}{HTML}{E09F3E}
\definecolor{auburn}{HTML}{9E2A2B}
\definecolor{rosswood}{HTML}{540B0E}
\definecolor{limegreen}{HTML}{32CD32}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \psi^*
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\draw[thick,->, opacity=1] (-1.4,0) -- (1.4,0) node[anchor= west, opacity=1]{\scriptsize $u_1$};
\draw[thick,->, opacity=1] (0,-1.4) -- (0,1.4)node[anchor=south, opacity=1]{\scriptsize $u_2$};

\draw[thick](0,.5) -- (.25,.75) -- (.25,.25) -- (.75,.25) -- (0.5,0) -- (.75,-.25) -- (.25,-.25) -- (.25,-.75)-- (0,-.5) 
-- (-.25,-.75) -- (-.25,.-.25) -- (-.75,-.25) -- (-.5,0) -- (-.75,.25)
-- (-.25,.25) -- (-.25,.75) -- (0,.5);

\draw[thick, fill=limegreen, opacity = 0.8](0,.5) -- (.25,.75) -- (.25,.25) -- (.75,.25) -- (0.5,0) -- (.75,-.25) -- (.25,-.25) -- (.25,-.75)-- (0,-.5) 
-- (-.25,-.75) -- (-.25,.-.25) -- (-.75,-.25) -- (-.5,0) -- (-.75,.25)
-- (-.25,.25) -- (-.25,.75) -- (0,.5);

\draw[fill=softyellow, opacity = 0.2,draw=none](.25,.25) -- (1.2,.25) -- (1.2,1.2) -- (.25,1.2) -- cycle;
\draw[thick](.25,1.2) -- (.25,.25) -- (1.2,.25);

\draw[draw=none,fill=softyellow, opacity = 0.2](-.25,-.25) -- (-1.2,-.25) -- (-1.2,-1.2) -- (-.25,-1.2) -- cycle;
\draw[thick](-.25,-1.2) -- (-.25,-.25) -- (-1.2,-.25);

\draw[draw=none,thick,fill=softyellow, opacity = 0.2](.25,-.25) -- (1.2,-.25) -- (1.2,-1.2) -- (.25,-1.2) -- cycle;
\draw[thick](.25,-1.2) -- (.25,-.25) -- (1.2,-.25);

\draw[draw=none,thick,fill=softyellow, opacity = 0.2](-.25,.25) -- (-1.2,.25) -- (-1.2,1.2) -- (-.25,1.2) -- cycle;
\draw[thick](-.25,1.2) -- (-.25,.25) -- (-1.2,.25);

\draw[draw=none,fill=limegreen, opacity = 0.2](-.5,0) -- (-.75,-.25)  -- (-1.2,-.25) -- (-1.2,.25) -- (-.75,.25) -- cycle;
\draw[thick] (-1.2,.25) -- (-.75,.25) -- (-.5,0) -- (-.75,-.25)  -- (-1.2,-.25);

\draw[draw=none,fill=limegreen, opacity = 0.2](.5,0) -- (.75,.25)  -- (1.2,.25) -- (1.2,-.25) -- (.75,-.25) -- cycle;
\draw[thick](1.2,-.25) -- (.75,-.25) -- (.5,0) -- (.75,.25)  -- (1.2,.25);

\draw[draw=none,fill=limegreen, opacity = 0.2](0,.5) -- (.25,.75)  -- (.25,1.2) -- (-.25,1.2) -- (-.25,.75) -- cycle;
\draw[thick](-.25,1.2) -- (-.25,.75) -- (0,.5) -- (.25,.75)  -- (.25,1.2);

\draw[draw=none,fill=limegreen, opacity = 0.2](0,-.5) -- (-.25,-.75)  -- (-.25,-1.2) -- (.25,-1.2) -- (.25,-.75) -- cycle;
\draw[thick](.25,-1.2) -- (.25,-.75) -- (0,-.5) -- (-.25,-.75)  -- (-.25,-1.2);

\node at (1,1){\textbullet} (1,1)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,1)$};
\node at (1,0){\textbullet} (1,0)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,0)$};
\node at (0,1){\textbullet} (-0.03,1)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,1)$};

\node at (-1,-1){\textbullet} (-1,-1)node[anchor=south, minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,-1)$};
\node at (-1,0){\textbullet} (-1,0)node[anchor= south west, minimum height = 0.7cm, scale = 0.4]{\tiny $(-1,0)$};
\node at (-1,1){\textbullet} (-1,1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,1)$};

\node at (0,-1){\textbullet} (-0.025,-1)node[anchor=south west,  minimum height = 0.75cm, scale = 0.3]{\tiny $(0, -1)$};
\node at (1,-1){\textbullet} (1,-1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1, -1)$};

\node at (0,0){\textbullet} (0,0)node[anchor=south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,0)$};

\end{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \psi^$\diamond$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\draw[thick,->, opacity=1] (-1.4,0) -- (1.4,0) node[anchor=west, opacity=1]{\scriptsize $u_1$};
\draw[thick,->, opacity=1] (0,-1.4) -- (0,1.4)node[anchor=south, opacity=1]{\scriptsize $u_2$};

\draw[thick](.75,.75) -- (.75,.-.75) -- (-.75,-.75) -- (-.75,.75) -- cycle;
\draw[thick, fill=limegreen, opacity = 0.8](.75,.75) -- (.75,.-.75) -- (-.75,-.75) -- (-.75,.75) -- cycle;

\draw[draw=none,fill=softyellow, opacity = 0.2](.75,.75) -- (.75,.25) --
(1,.5)--(1.2,.5) -- (1.2,1.2) -- (.5,1.2) -- (.5,1) -- (.7,1.2) -- (.25,.75) --cycle;
\draw[thick](.5,1.2) -- (.5,1)--(.25,.75)--(.75,.75) -- (.75,.25) --
(1,.5)--(1.2,.5);

\draw[draw=none,fill=softyellow, opacity = 0.2](-.75,-.75) -- (-.75,-.25) -- (-1,-.5) -- (-1.2,-.5) -- (-1.2,-1.2) -- (-.5,-1.2)-- (-.5,-1) -- (-.25,-.75) --cycle;
\draw[thick](-.5,-1.2) -- (-.5,-1)--(-.25,-.75)--(-.75,-.75) -- (-.75,-.25) --
(-1,-.5)--(-1.2,-.5);

\draw[draw=none,fill=softyellow, opacity = 0.2](.75,-.75) -- (.75,-.25) -- (1,-.5) -- (1.2,-.5) -- (1.2,-1.2) -- (.5,-1.2) --(.5,-1) -- (.25,-.75) --cycle;
\draw[thick](.5,-1.2)--(.5,-1)--(.25,-.75)--(.75,-.75) -- (.75,-.25) -- (1,-.5) -- (1.2,-.5);

\draw[draw=none,fill=softyellow, opacity = 0.2](-.75,.75) -- (-.75,.25) --  (-1,.5)--(-1.2,.5) -- (-1.2,1.2) -- (-.5,1.2) --(-.5,1) -- (-.25,.75) --cycle;
\draw[thick](-.5,1.2)--(-.5,1)--(-.25,.75)--(-.75,.75) -- (-.75,.25) -- (-1,.5) -- (-1.2,.5);

\draw[draw=none,fill=limegreen, opacity = 0.2](0,1.2) -- (.5,1.2)--(.5,1)  -- (.25,.75) -- (-.25,.75) -- (-.5,1)--(-.5,1.2)  -- cycle;
\draw[thick](-.5,1.2)--(-.5,1) -- (-.25,.75) -- (.25,.75) -- (.5,1) -- (.5,1.2);

\draw[draw=none,fill=limegreen, opacity = 0.2](-1.2,-.5) --(-1,-.5) -- (-.75,-.25)  -- (-.75,.25) -- (-1,.5)  --(-1.2,.5)-- cycle;
\draw[thick](.5,-1.2)--(.5,-1) -- (.25,-.75) -- (-.25,-.75) -- (-.5,-1) -- (-.5,-1.2);

\draw[draw=none,fill=limegreen, opacity = 0.2](1.2,-.5)--(1,-.5) -- (.75,-.25)  -- (.75,.25) -- (1,.5)--(1.2,.5)  -- cycle;
\draw[thick](-.5,-1.2)--(-.5,-1)--(-.25,-.75)--(-.75,-.75) -- (-.75,-.25) -- (-1,-.5) -- (-1.2,-.5);

\draw[draw=none,fill=limegreen, opacity = 0.2](0,-1.2) --(.5,-1.2)-- (.5,-1)  -- (.25,-.75) -- (-.25,-.75) -- (-.5,-1) -- (-.5,-1.2)  -- cycle;
\draw[thick](.5,-1.2)--(.5,-1)--(.25,-.75) -- (-.25,-.75) -- (-.5,-1) -- (-.5,-1.2);

%\path[fill=yellow, opacity = 0.1] (1.2,1.2) -- (1.2,-1.2) -- (-1.2,-1.2) -- (-1.2,1.2)  -- cycle;

\node at (1,1){\textbullet} (1,1)node[anchor= south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,1)$};
\node at (1,0){\textbullet} (1,0)node[anchor= south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,0)$};
\node at (0,1){\textbullet} (-0.03,1)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,1)$};

\node at (-1,-1){\textbullet} (-1,-1)node[anchor=south, minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,-1)$};
\node at (-1,0){\textbullet} (-1,0)node[anchor= south, minimum height = 0.7cm, scale = 0.4]{\tiny $(-1,0)$};
\node at (-1,1){\textbullet} (-1,1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,1)$};

\node at (0,-1){\textbullet} (-0.025,-1)node[anchor=south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0, -1)$};
\node at (1,-1){\textbullet} (1,-1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1, -1)$};

\node at (0,0){\textbullet} (0,0)node[anchor=south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,0)$};

\end{tikzpicture}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: