### 2022-06-29 poly-embed-journal/jmlr_cover_letter/jmlr_coverletter.tex
Submitted arXiv version, and converted the original to JMLR format.  Added cover letter

% !TEX program = pdflatex
\input{jmlr_coverletter_preamble.tex}

\begin{document}
\hfill{\today}

Journal of Machine Learning Research

\bigskip

Dear Editors:

We are writing to submit our manuscript ``An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates'' to the \emph{Journal of Machine Learning Research}.

Our manuscript studies \red{the design of consistent piecewise linear and convex surrogates for general discrete prediction tasks} which is important because \red{this is the first framework to our knowledge that characterizes consistent and convex surrogates for general prediction tasks.  Convex surrogates for prediction tasks are often design in an ad-hoc manner, and as a result are often not actually consistent for the prediction task at hand.}

This paper extends our works of \red{An Embedding Framework for Consistent Polyhedral Surrogates} and \red{Surrogate Regret Bounds for Polyhedral Losses} by \red{adding results connecting our proposed embedding framework to indirect elicitation, adding more examples and giving new results about their surrogate regret bounds, and fully establishing when a convex surrogate embeds a discrete loss, among others}.

We suggest the following action editors and referees for our submission.

Action Editors:
\begin{itemize}
    \item \red{Sanmi Koyejo, University of Illinois at Urbana-Champaign (sanmi@illinois.edu)}
    \item \red{Ambuj Tewari, University of Michigan (tewaria@umich.edu)}
    \item \red{Moritz Hardt, Max Planck Institute for Intelligent Systems, T\"ubingen ( hardt@is.mpg.de)}
    \item \red{Tong Zhang,  The Hong Kong University of Science and Technology ( tongzhang@tongzhang-ml.org)}
\end{itemize}

Reviewers:
\begin{itemize}
    \item \red{Bob Williamson, University of T\"ubingen (bob.williamson @uni-tuebingen.de )}
    \item \red{Shivani Agarwal, University of Pennsylviania (ashivani@seas.upenn.edu)}
    \item \red{Alex Nowak-Vila, Owkin (alex.nowak-vila@inria.fr)}
    \item \red{First Lastname, Institution (first.lastname@institution.edu)}
    \item \red{First Lastname, Institution (first.lastname@institution.edu)}
\end{itemize}

Our submission has the following keywords: \red{Surrogate loss function}, \red{Statistical (FISHER? UNIVERSAL?) consistency}, \red{keyword}, \red{keyword}, \red{keyword}.

As the corresponding author, I confirm that none of the co-authors listed below have a conflict of interest with the action editors and referees I suggest above. Further, I confirm that all co-authors below consent to my submission of this manuscript to the \emph{Journal of Machine Learning Research}.

\bigskip

Sincerely,

\medskip

Jessie Finocchiaro (University of Colorado Boulder)\\
Rafael Frongillo (University of Colorado Boulder)\\
Bo Waggoner (University of Colorado Boulder)

\end{document}