### 2022-06-29 poly-embed-journal/jmlr_cover_letter/jmlr_coverletter.tex
cover letter pass

% !TEX program = pdflatex
\input{jmlr_coverletter_preamble.tex}

\begin{document}
\hfill{\today}

Journal of Machine Learning Research

\bigskip

Dear Editors:

We are writing to submit our manuscript ``An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates'' to the \emph{Journal of Machine Learning Research}.

Our manuscript proposes a new framework to design consistent convex surrogates for general discrete prediction tasks.
As convex surrogates are often designed in an ad-hoc manner, and often are inconsistent for the desired target task, our framework fills a urgent gap in the literature.

Our manuscript extends the conference publication \emph{An Embedding Framework for Consistent Polyhedral Surrogates} (NeurIPS 2019) and incorporates roughly half the results from \emph{Surrogate Regret Bounds for Polyhedral Losses} (NeurIPS 2021).
We have substantially rewritten and reorganized these papers, as well as adding entirely new results (\S 6, \S 7, \S E).
For example, \S 6 shows the equivalence of several natural notions of non-redundancy, and \S 7 uses our framework to show that indirect elicitation is equivalent to consistency for polyhedral surrogates.

We suggest the following action editors, referees, and keywords for our submission.

Action Editors:
\begin{itemize}
    \item Ambuj Tewari, University of Michigan (\url{tewaria@umich.edu})
    \item Sanmi Koyejo, University of Illinois at Urbana-Champaign (\url{sanmi@illinois.edu})
    \item Moritz Hardt, Max Planck Institute for Intelligent Systems, T\"ubingen (\url{hardt@is.mpg.de})
    \item Tong Zhang,  The Hong Kong University of Science and Technology (\url{tongzhang@tongzhang-ml.org})
\end{itemize}

Reviewers:
\begin{itemize}
    \item Bob Williamson, University of T\"ubingen (\url{bob.williamson@uni-tuebingen.de})
    \item Shivani Agarwal, University of Pennsylviania (\url{ashivani@seas.upenn.edu})
    \item Alex Nowak-Vila, Owkin (\url{alex.nowak-vila@inria.fr})
    \item First Lastname, Institution (first.lastname@institution.edu)
    \item First Lastname, Institution (first.lastname@institution.edu)
\end{itemize}

Keywords: Statistical consistency, surrogate loss functions, calibration, property elicitation, embeddings.

As the corresponding author, I confirm that none of the co-authors listed below have a conflict of interest with the action editors and referees suggested above. Further, I confirm that all co-authors below consent to my submission of this manuscript to the \emph{Journal of Machine Learning Research}.

\bigskip

Sincerely,

\medskip

Jessie Finocchiaro (University of Colorado Boulder)\\
Rafael Frongillo (University of Colorado Boulder)\\
Bo Waggoner (University of Colorado Boulder)

\end{document}